
---
title: "Functions"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(tidyverse)
library(gradethis)
library(PPBDS.data)
library(learnr)
library(purrr)
tutorial_options(exercise.checker = gradethis::grade_learnr)
knitr::opts_chunk$set(echo = FALSE, exercise.checker = grade_learnr)
options(tutorial.storage="client") 

```



## Functions - From RPrimer (1)

###  

Functions are the verbs of R. When you do something in R, you do it by running a function. For example,

* To take a log, you run a function named log: `log(2.71)`.
* To fit a linear model, you run a function named lm: `lm(mpg ~ wt, data = mtcars)`.
* To add two numbers, you run a function named "+": `1 + 2`. 

This last one is an alternative way to call the function `'+'(1, 2)`. In R, even things that do not look like functions are functions. 

###  

But what is a function? 

A function is simply a piece of code that is packaged in a way that makes it easy to reuse.

This short tutorial will show you how R functions are packaged and how you can work with R functions. It is a pre-requisite for the next tutorial, which will show you how to write your own R functions. Let's begin!

### Parts of a function

###  

Every R function has three parts:

1. A body of code
1. A set of formal arguments
1. An environment (where the function will look up the values of the objects within it)

###  

You can inspect each part of a function with three helper functions:

1. `body()`
1. `formals()`
1. `environment()`

Let's use these helpers to see how the `xor` function works. `xor` is a logical operator that returns `TRUE` if one, but not both, or its arguments evaluates to `TRUE`.

### `body()`

* Run `body(xor)` to see the code body of `xor`. Then click Submit Code.

```{r xor-body, exercise = TRUE, echo = TRUE}

```

```{r xor-body-solution}
body(xor)
```

```{r xor-body-check}
grade_code("`xor` contains one line of code in its body (you can ignore the braces). This isn't always the case. A function can have as many lines of code as it needs in its body. If you don't believe me check out the body for lm. R will run the code in a function's code body each time you call the function; but as you can see, the code in a function's body may not be self-contained. What is x? What is y?")
```

###  

`xor()` contains one line of code in its body (you can ignore the braces). This isn't always the case. A function can have as many lines of code as it needs in its body. For example, `as.list.data.frame` has three lines (most R functions have much more).

```{r, echo = TRUE}
body(as.list.data.frame)
```

When a function contains more than one line of code in its body, R will run the entire body and return the result of the **last** line.

###  

R will run the code in a function's code body each time you call the function; but as you can see with `xor`, the code in a function's body may not be self-contained. 

```{r, echo = TRUE}
xor
```

What is `x`? What is `y`? Let's find out.

### `formals()`

Often times, a function will need one or more named values to do its job. These values are the _formal arguments_ of the function. The formal arguments of `xor` are `x` and `y`.

`formals()` returns a list of the formal arguments of a function. If an argument in the list does not come with a value, you will need to supply a value when you call the function. If a function does not have any formal arguments (always possible), `formals()` will return `NULL`.

* Use `formals()` to see the formal arguments of `xor` (hint: they are `x` and `y`). Then click Submit Answer.

```{r xor-formals, exercise = TRUE, eval = TRUE}

```

```{r xor-formals-solution, echo = FALSE}
formals(xor)
```

```{r xor-formals-check}
grade_code(A function can have zero, one, or many many formal arguments. Do not be distracted by the $'s that appear before the argument names. Those are an artifact of how R displays lists.)
```

### `environment()`

Each function is associated with an R environment. The association tells R how to look up any objects that are used by a function, but not defined in its arguments. We'll learn more about environments later, so you do not need to worry about them now. However, looking up a function's environment is a handy way to tell which package the function comes in.

* Use `environment()` to see the environment associated with the `gather` function.

```{r gather-env, exercise = TRUE}

```

```{r gather-env-solution}
environment(gather)
```

```{r gather-env-check}
grade_code("The `gather()` function comes in the tidyr package. If a function's environment is namespace:base, that is a good sign that the function comes in base R, the part of R that loads before you load any optional packages.")
```

### names

You can look up all three parts of a function at once by typing the name of the object that the function is stored in. This is what you think of as the "name of the function." (In reality, functions do not have names, but they are stored in objects that have names.)

For example, the `xor` function is a nameless function stored in the object named `xor`. To see the function in full, type `xor`.

```{r name, exercise = TRUE}
xor
```

### Help pages

And _never_ forget: you can learn more about any R function by looking up its help page. To do this, type the "function's name" preceded by a question mark, e.g.

```{r eval = FALSE, echo = TRUE}
?xor
```

When you type this into an R IDE, such as the R console, a help page will open in a separate tab.

### Quiz

```{r q1, echo = FALSE}
question("Which of the following is not a part of a function?",
         allow_retry = TRUE,
         random_answer_order = TRUE,
         answer("A name", correct = TRUE, message = "That's right, functions contain a code body, a set of formal arguments, and an environment, but not a name. They just happen to inherit a name from the object they are stored in."),
         answer("A code body"),
         answer("A (potentially empty) set of formal arguments"),
         answer("An environment"))
```

### Primitives

You should also know about an important class of R functions called _primitives_. Most R functions call other R functions in their code body, but primitives do not. Instead primitives call internally implemented algorithms that are written in a lower level programming language (i.e. a more primitive programming language) like C, C++, or FORTRAN. 

R's primitive functions are designed to be very fast, but you won't be able to inspect their components. See what happens if you inspect the contents of `log`, which is a primitive function.

```{r log, exercise = TRUE, eval = FALSE}
body(log)
formals(log)
environment(log)
```

That's enough minutiae! Now, how do you _run_ a function?

### Calling functions

###  

To call an R function, type its name followed immediately by an open parenthesis and eventually by a closed parenthesis. The parentheses act as a "trigger" that causes R to run the code in the function's code body. Note the difference:

```{r sysdate, exercise = TRUE, eval = FALSE}
Sys.Date
Sys.Date()
```

If you ever try to run a function and nothing happens, double check to see if you included the closing parenthesis.

###  

If a function requires you to provide values for formal arguments, provide the values between the parentheses. 

```{r xoragain, exercise = TRUE, eval = FALSE}
xor(x = TRUE, y = FALSE)
```

The best practice is to explicitly write out each argument name followed by an equals sign and then a value. If you leave out the argument names, R will match your values to arguments in the order that they are listed. It is common practice among R users to skip the first one or two argument names and then write out the rest.

### Default values

Some R functions will provide a default value for one or more of arguments. In this case, you do not need to pass a value to the argument when you call the function. If you do not provide a value, the function will use the default value. 

You will know that an argument comes with a default value, if the argument appears with a value in the `formals()` list.

* `rnorm` generates n random values from a normal distribution. When you call `rnorm` you can supply the mean and standard deviation of the distribution to draw the values from. Check to see if `rnorm` uses any default values. For which arguments? What are the values? Click Submit Answer to see if you are right.

```{r rnorm, exercise = TRUE}

```

```{r rnorm-solution, echo = FALSE}
formals(rnorm)
```

```{r rnorm-check}
grade_code("rnorm provides a default value of zero for the mean argument and a default argument of one for the sd argument. In other words, if you do not supply a mean and sd value when you call rnorm, rnorm will randomly generate results from a standard normal distribution. Default values can be very handy!")
```

###  

You can also spot a function's default values by examining the function itself. Run the code below to see them. 

```{r rnormname, exercise = TRUE}
rnorm
```

###  

There is one last thing you should know about functions: functions are like Las Vegas:

> What happens in a function stays in the function.
For example, suppose I've saved values as `x` and `y`. 

```{r tryout-setup}
x <- 1
y <- FALSE
```

###  

Calling `xor` will not change those values, even though I'm defining new values for `x` and `y` in the function.

```{r tryout, exercise = TRUE}
xor(x = TRUE, y = TRUE)
x
y
```

###   

Moreover, `xor` won't remember which values I've used in the past. If I forget to define `x` and `y` the next time I call `xor`, `xor` will return an error. `xor` won't even be able to look up the values that I assigned to `x` and `y` outside of the function.

```{r ex4, exercise = TRUE}
xor()
```

_What happens in a function stays in the function._ 

So don't worry about using functions. (Ahem, technically you _could_ write a function that would change things outside of the function, but that would be a bad idea, and you would have to try to do it on purpose, and it would be rather hard at that. So don't worry.)

### Quiz

#### Exercise 1

```{r ex1, echo = FALSE}
question("Which of these will run the `Sys.time` function?",
         answer("`Sys.time`"),
         answer("`Sys.time()`", correct = TRUE, 
         message = "The open and closed parentheses tell R to execute the code in the code body of the function stored in the object names Sys.time."))
```

#### Exercise 2

* Find the code body of `foo`


```{r foo-setup, echo = FALSE, eval = TRUE}
foo <- function() {
  a <- 10
  a
}
```

```{r foo, exercise = TRUE}

```


```{r foo-solution, echo = FALSE}
body(foo)
```

```{r foo-check}
grade_code("Good!")
```

#### Exercise 3

Recall the code in `foo`:

```{r echo = FALSE}
foo <- function() {
  a <- 10
  a
}
```

```{r, echo = TRUE}
body(foo) 
```

What will be the final value of `a` if I run the following three lines of code in order?

```{r eval = FALSE, echo=TRUE}
a <- 1
foo()
a
```

```{r ex3, echo = FALSE}
question("",
         answer("1", correct = TRUE, message = "What happens in a function, stays in the function. `foo()` will not change the value of a outside of `foo()`."),
         answer("10"),
         answer("Running a will return an error."),
         allow_retry = TRUE)
```

#### Congratulations!

You have finished the tutorial.


*Let's start off by recapping what we learnt at the beginning of our section on writing functions. Recall that we used a variety of different ways to calculate the same output - the range of a numeric variable*


## Building your first function 

* Let's start off by recapping what we learnt at the beginning of the chapter on Functions. Recall that we used a variety of different ways to calculate the same output - the range of a numeric variable*


*Let's start off by recapping what we learnt at the beginning of the chapter on Functions. Recall that we used a variety of different ways to calculate the same output - the range of a numeric variable*


```{r quiz-1}



quiz(
  question("Which of the following commands yield the correct range (3293282) of the adult population in the Midwest data?",
             answer("`max(midwest$popadults) - min(midwest$popadults)`"),
             answer("`range(midwest$popadults)[2] - range(midwest$popadults)[1]`"),
             answer("`diff(range(midwest$popadults))`"),
             answer("All of the above.", correct = TRUE),
             allow_retry = TRUE
    )
)



```

*Test the code here to verify that you get the same output range:*

```{r exercise-1, exercise = TRUE, exercise.lines = 5}
---------

---------
```


### 

*We will now use code from the previous question to write our first, simple R function. Use the code `max() - min()` to write your first, simple R function to calculate the range of the adult population:*

```{r exercise-2, exercise = TRUE, exercise.lines = 10}

# Write the function by using `max() - min()`
# from the previous question. 

range_function <- function(x) --------
  
# Now call the range_function to calculate the range
# of the adult population in the Midwest dataset.
  
range_function(------)

```

```{r exercise-2-solution}

range_function <- function(x) max(x) - min(x)

range_function(midwest$popadults)
```

```{r exercise-2-check}
grade_code()
```

*Congratulations! You've written your first, simple R function! Note that this function can apply to _any_ numeric variable, not just the `popadults` variable in the midwest dataset. We say that this function is generalized.*

```{r, quiz-2}

quiz(
  question("What is not a benefit of writing a function?",
             answer("Your code is reusable."),
             answer("You make your analysis and code harder to understand.", correct = TRUE),
             answer("You save time and energy carrying out the same processes repeatedly."),
             answer("None of the above."),
             allow_retry = TRUE
    ),
  question("Suppose you need to run the following code on 20 different objects, `sqrt(sum(x^2))`. Why would you not want to rewrite the code 20 times? (Check all that apply)",
     answer("Retyping code takes time.", correct = TRUE),
     answer("Typing the code 20 times creates 20 chances to make a typo.", correct = TRUE),
     answer("If you need to change the code later, you'll need to change it in 20 different places.", correct = TRUE),
     allow_retry = TRUE, 
     type = "multiple", 
     incorrect = "Did you check ALL that applied?",
     correct = "Good job! You can avoid these outcomes by turning your code into a function. A good rule is to turn a piece of code into a function whenever you find yourself re-using the code more than three times.")
  
)
  
```

## How to write a function (R Primer 2)

###  When to write a function

R comes with thousands of functions, and more are written everyday (these are published in R packages). If you want to do something in R, there is a good chance that a function already exists somewhere that does it.

But what if you want to do something new in R, something that doesn't yet have a function? What should you do?

###  

Write a function?

No, silly. Write some code. But what if you want to use that code _over and over again_, as if it were a part of R? Then what should you do?

###  

Write a function!

This tutorial will show you how. Take the quiz below to begin.

### Quiz

Suppose you need to run the following code on 20 different objects, `sqrt(sum(x^2))`.

```{r why-quiz, echo = FALSE}
question("Why would you not want to rewrite the code 20 times? (Check all that apply)",
     answer("Retyping code takes time.", correct = TRUE),
     answer("Typing the code 20 times creates 20 chances to make a typo.", correct = TRUE),
     answer("If you need to change the code later, you'll need to change it in 20 different places.", correct = TRUE),
     allow_retry = TRUE, 
     type = "multiple", 
     incorrect = "Did you check ALL that applied?",
     correct = "Good job! You can avoid these outcomes by turning your code into a function. A good rule is to turn a piece of code into a function whenever you find yourself re-using the code more than three times.")
```


### Workflow

In the previous section, we created a very simple function used to calculate the range of a numeric variable. In general, however, you must always follow these four steps to turn any piece of code into a function: 

1. Create a real R object (or set of objects) to use with your function
2. Write code that works with the real object(s)
3. Wrap the code in `function()`
4. Assign the names of your real objects as argument names to the function

These steps are the best practice for writing functions in R. When you follow these steps, your functions are guaranteed to work for at least one case, which means you will spend more time using your functions and less time debugging them.

Let's use the steps to create your first function.

### Goal - Grading

To make this real, put yourself in the shoes of a teacher: 

You've given your students 10 homework assignments and announced that you will drop their lowest homework score. Their final grade will be the average score of the remaining homeworks.

To make your life easier, you want to write an R function that will take a vector of 10 homework scores and return a final grade. 

Ready to begin?

###  

Remember our steps:

1. Create a real R object (or set of objects) to use with your function

* Create an object named `x` that contains the vector `c(100, 100, 100, 100, 100, 100, 100, 100, 100, 90)` (hint: use copy and paste!). Then click Submit Answer. `x` will be the grades of your test student. 

```{r obj1, exercise = TRUE}
x <- ------------------------------------

```

```{r obj1-solution}
x <- c(100, 100, 100, 100, 100, 100, 100, 100, 100, 90)
```

```{r obj1-check}
"Now you can use x to develop and test your code"
```

###  

2. Write code that works with the real object(s)

Recall the grading scheme: the final grade is the average homework score after you drop the lowest score. Since there are many ways to calculate this, let's be specific and end up on the same page.

* Use `sum()`, `min()`, `/`, `9` and parentheses to calculate the final grade for student `x`. Click Submit Answer to check that your code works.

```{r make-x, echo = FALSE}
x <- c(100, 100, 100, 100, 100, 100, 100, 100, 100, 90)
```

```{r norm, exercise = TRUE, setup = "make-x"}
x
```

```{r norm-hint}
"Your code should return 100."
```

```{r norm-solution}
(sum(x) - min(x)) / 9
```

```{r norm-check}
"This is the crucial step in writing a function. It is important to get your code working in real life with a real object before you try to abstract the code into a reusable function."
```

###  

3. Wrap the code in `function()`

The `function()` function builds a function from a piece of R code. To use it, call `function()` followed by an opening brace, `{`. Then write one or more lines of R code followed by a closed brace, `}`, e.g.

```{r, echo = TRUE}
foo <- function() {
 a <- 1
 b <- 2
 a + b
}
```

`function()` will return a function that uses everything between the braces as its code body. If you'd like to save the function, you'll need to assign it in the usual way to an R object that you can call later.

As you write your functions, recall that R will only return the result of the last line in the code body when you call the function (we'll learn about some exceptions to this rule in the Control Flow tutorial). 

###  

Once you save a function, you can run it and inspect its contents.

```{r, echo = TRUE}
foo()
foo
```

### Your turn

Let's save your code as a function.

* Save the code below as a function named `grade`. Then click Submit Answer.

```{r grade, exercise = TRUE}
(sum(x) - min(x)) / 9
```

```{r grade-hint}
"Use the function() {} function."
```

```{r grade-solution}
grade <- function() { 
  (sum(x) - min(x)) / 9
}
```

```{r grade-check}
"Now that you've saved grade as a function, you can call it whenever you like, but there is still one more thing to do..."
```

###  

At the moment, your `grade()` function is reusable but not _generalizable_. Each time you call `grade()` it computes the final grade of the vector `x` that contains `c(100, 100, 100, 100, 100, 100, 100, 100, 100, 90)`.

```{r echo = FALSE}
x <- c(100, 100, 100, 100, 100, 100, 100, 100, 100, 90)
grade <- function() { 
  (sum(x) - min(x)) / 9
}
```

```{r}
grade()
grade()
```

We'd like to use `grade()` with new vectors that have new values.

### Arguments

4. Assign the names of your real objects as argument names to the function

You can make a function generalizable by turning some of the objects in its code body into _formal arguments_. A formal argument is an object that a user can assign a value to when he or she calls the function. The function will use the user's value for the object when it executes its code body.

For example, we'd like to tell R that `x` in `grade()` is an argument. R shouldn't use a pre-defined value for `x`; it should let the user supply a new value for `x` each time he or she runs the function.

```{r eval = FALSE}
grade <- function() { 
  (sum(x) - min(x)) / 9
}
```

###  

How do you tell R that an object is a formal argument? 

You list the name of the object in the parentheses that follow `function()` in the function definition. If you make more than one argument, separate their names with a comma. For example, you could make `a` and `b` arguments of my `foo` function.

```{r}
foo <- function(a, b) {
 a + b
}
```

Now I can define a new value for `a` and `b` each time I call `foo`. 

```{r, echo = TRUE, eval = TRUE}
foo(a = 1, b = 1)
foo(a = 100, b = 200)
```

### Default values

To give an argument a default value, set it equal to a value when you define the function. For example, the code below will set the default value of `b` in `foo` to one.

```{r, echo = TRUE, eval = TRUE}
foo <- function(a, b = 1) {
 a + b
}
foo(a = 2)
foo(a = 2, b = 2)
```

###  

Interesting, huh? Now apply what you've learned to `grade()`.

* Change the code below to list `x` as a formal argument of `grade()`. Then click Submit Answer.

```{r x, exercise = TRUE}
grade <- function() { 
  (sum(x) - min(x)) / 9
}
```

```{r x-solution}
grade <- function(x) { 
  (sum(x) - min(x)) / 9
}
```

```{r x-check}
'If you use the "best practice" workflow, your argument names will always be the names of the real R objects that you create in Step 1. Nice and simple.'
```

###   

Good job! You can now have a finished `grade()` function that you can use to calculate the final grade of _any_ vector (I mean, student). Try it out.

* Calculate the final grade of the vector `c(100, 90, 90, 90, 90, 90, 90, 90, 90, 80)`. Then click Submit Answer.

```{r make-grade}
grade <- function(x) { 
  (sum(x) - min(x)) / 9
}
```

```{r function, exercise = TRUE, exercise.setup = "make-grade"}
grade(x = ------------------------------------------)
```

```{r function-solution}
grade(x = c(100, 90, 90, 90, 90, 90, 90, 90, 90, 80))
```


```{r function-check}
grade_code(correct = "Let's recap the function writing workflow.")

```

### Quiz

Congratulations on finishing the tutorial. Before you go, complete the following exercises to cement your skills!

#### Exercise 1

```{r e1, echo = FALSE}

question("What is the first step of writing a function?",
         answer("Choose a name, then type `<- function(){`", message = "You shouldn't type `function()` until you have code that works with a real test case."),
         answer("Create a real R object (or set of objects)", correct = TRUE, message = "You can then use the object to test your code."),
         answer("Assign argument names to your function"),
         answer("Write a help page that describes how your function will work."),
         allow_retry = TRUE)
```

#### Exercise 2

What will the following code return?

```{r eval = FALSE, echo = TRUE}
f <- function(a, b, c) {
  a
  b
  c
}
f(1, 2, 3)
```

```{r q2, echo = FALSE}
question("",
         answer("1"),
         answer("2"),
         answer("3", correct =TRUE, message = "A function will return the result of its *last* line."),
         answer("A vector that contains 1, 2, and 3"),
         answer("An error"),
         allow_retry = TRUE)
```

#### Exercise 3

Let's write a function that grades by counting a student's highest score twice and then taking an average of the 11 scores.

```{r e2, echo = FALSE}
question("Do you have a test vector that you can use?",
         answer("Test vector? We don't need no stinking test vector.", 
                message = "Yes you do."),
         answer("Yes, we can reuse the vector named x.", correct = TRUE),
         allow_retry = TRUE,
         random_answer_order = TRUE)
```

#### Exercise 4

* Write a piece of code that uses `max()` to double count the highest score in `x` and then takes the average of the 11 results. Click Submit Answer to check that your code works.

```{r e3, exercise = TRUE, exercise.setup = "make-x"}
x
```

```{r e3-hint}
"Your code should retrun the answer 99.09091."
```

```{r e3-solution}
(sum(x) + max(x)) / 11
```

```{r e3-check}
"That student will be pleased!"
```

#### Exercise 5

* Wrap your code in `function()` and save it to the name `grade2`.
* Then define the argument(s) for the function.
* Then click Submit Answer.

```{r e4, exercise = TRUE}
(sum(x) + max(x)) / 11
```

```{r e4-hint}
"When you use this workflow, the names of the argument(s) are always the names of the object(s) that you created in step 1."
```

```{r e4-solution}
grade2 <- function(x) {
  (sum(x) + max(x)) / 11
}
```

```{r e4-check}
"Your skills are almost complete."
```

#### Exercise 6

It's time to put it all together. 

* Use the code block below to develop a function named `l2` that calculates the Euclidean distance, or L~2~ norm ($\|x||_{2}$), of a vector of numbers named `x`. The L~2~ norm is the square root of the sum of the squared values in the vector, i.e.

$$\| x \|_{2} = \sqrt{\sum_{i = 1}^{n} x_{i}^{2}}$$

* When you are finished, reduce your code to just the definition of your function. Then click Submit Answer.

```{r e5, exercise = TRUE}
l2 <- 

```

```{r e5-hint-1}
"You've already seen the code for the l2 function earlier in this tutorial (I do not mean the pipe example)."
```

```{r e5-hint-2}
"Consider using sqrt(), sum(), and ^2. Your code will work with vectors of any length."

```


```{r e5-solution}
l2 <- function(x) {
  sqrt(sum(x^2))
}
```

```{r e5-check}
"Good job!"
```



## Test your function

### Quiz:

Let's go back to our `range_function` for a bit, and test it with varying inputs. Remember, we built the range_function to help us calculate the range of a numeric variable, popadults.

```{r, quiz-3}

quiz(
  question("If we ran the following lines of code, which of the options will not yield errors?",
             answer("`range_function('I like apples more than bananas')`"),
             answer("`range_function(midwest)`"),
             answer("`range_function(midwest$popwhite)`", correct = TRUE),
             answer("All of the above options will yield errors."),
             allow_retry = TRUE
    ),
  
  question("What happens if we run the following code: `range_function(c(FALSE, TRUE, TRUE, FALSE))`?",
             answer("R shuts down."),
             answer("R will give us an error explaining why the code cannot work."),
             answer("R will coerce the logical vectors to ones and zeroes, and caluclate their range.", correct = TRUE),
             answer("R will coerce the logical vectors to numeric vectors, and caluclate their range."),
             allow_retry = TRUE)
  
  
)
```

### Building a new function, and putting it to test: 

Now help us write a function that computes the sum of an input object, and multiplies it by 4. Call this function `sumtimesfour`.

```{r testfunction, exercise = TRUE}


sumtimesfour <- function(x) {
  
  -------
  
}
```

```{r testfunction-hint}
"Use `sum(x) * 4`"
```

```{r testfunction-solution}
sumtimesfour <- function(x) {
  
  sum(x) * 4
  
}

```

```{r testfunction-check}
grade_code()
```

Now, copy the following object onto your R chunk `n <- c(TRUE, FALSE, TRUE, FALSE)`. Then input `n` into the sumtimesfour function. 

```{r testfunction1, exercise = TRUE}


sumtimesfour <- function(x) {
   sum(x) * 4
}

n <- ----------
  
sumtimesfour(--)  
```

```{r testfunction1-solution}


sumtimesfour <- function(x) {
  sum(x) * 4
}

n <- c(TRUE, FALSE, TRUE, FALSE)

sumtimesfour(n)  

```


```{r testfunction1-check}
grade_code()
```

Recall that R tries to compute our input arguments, even though the outputs produced make absolutely no sense at all. This points us to another, incredibly important task - verifying our arguments, to ensure that they only compute values for the appropriate argument inputs. 

## Check the validity of arguments

```{r, quiz-6}

quiz(
  question("What is a function's argument?",
             answer("The line of logical reasoning used by a function to compute an output."),
             answer("Values used by a function to compute a result.", correct = TRUE),
             answer("Both."),
             answer("Neither."),
             allow_retry = TRUE
    ),
  
  question("Why is it a good idea to verify a function's argument? Check all that apply.",
             answer("To stop R from presenting us with non-sensical outputs.", correct = TRUE),
             answer("To display error messages every time an inappropriate value is used as a function's argument.", correct = TRUE),
             answer("To make sure R coerces the argument inputs to make sense of our requests."),
             answer("All of the above."),
             type = "multiple", 
             allow_retry = TRUE)
  
)

```

Let's now move on to create a new function, and verify its argument. Recall the Rule of Repair: 'When you must fail, fail noisily and as soon as possible'. We will try our best to act on this rule!

Let's compose a function that converts temperatures from the Celsius, to the Kelvin scale. The formula to do so is as follows: `celsius + 273.15`. Use this to develop your fourth R function of the day. We call this function `celsius_to_function`. 

```{r verifyfunction, exercise = TRUE}

# Copy the fomula code to compute the value of the temperature in Kelvins. 


celsius_to_kelvin <- function(celsius) {
  kelvin <- ------
  return(kelvin)
}

```

```{r verifyfunction-solution}


celsius_to_kelvin <- function(celsius) {
  kelvin <- celsius + 273.15
  return(kelvin)
}

```

```{r verifyfunction-check}
grade_code()
```

You've created yet another successful function. *Note that in R, it is not necessary to include the `return()` statement. R automatically returns whichever variable is on the last line of the body of the function. While in the learning phase, we will explicitly define the return statement. You absolutely must use return() if you want to return early based on some condition, i.e. before execution gets to the last line of the body. Otherwise, you can decide your own conventions about when you use return() and when you don’t.*

### Error messages with `stopifnot()`

Moving forward, we want to make sure that the input used by this function to convert from Celsius to Kelvin is a numeric vector. Temperatures are to only be recorded in numeric terms, for the purposes of this conversion.

One way to do this is to use the `stopifnot()` function. 

```{r verifyfunction1, exercise = TRUE}

# Use `stopifnot()` to verify that the input argument is a numeric vector. 
# Once more, use `celsius + 273.15` to compute the value of the temperature in Kelvins. 

celsius_to_kelvin <- function(celsius) {
  -------------
  kelvin <- ------
  return(kelvin)
}

```

```{r verifyfunction1-hint}
"Use `is.numeric(celsius)` wrapped in `stopifnot()`. Remember, we want to ensure that the `celsius` input is a numeric one."
```


```{r verifyfunction1-solution}


celsius_to_kelvin <- function(celsius) {
  stopifnot(is.numeric(celsius))
  kelvin <- celsius + 273.15
  return(kelvin)
}

```

```{r verifyfunction1-check}
grade_code()
```

Great! You have successfuly used `stopifnot()` as part of this function. Let's check if the function works with a non-numeric vector input. 

In the next chunk, create the object `test`. This must contain the vector `c(TRUE)`.

Next, input the object `test` into the `celsius_to_kelvin()` function. See if it works by running the code. 

```{r verifyfunction2, exercise = TRUE}

# Create vector `test` here.
-----------

# celsius_to_kelvin() function: 

celsius_to_kelvin <- function(celsius) {
  stopifnot(is.numeric(celsius))
  kelvin <- celsius + 273.15
  return(kelvin)
}

# Input `test` into the celsius_to_kelvin() function. 

celsius_to_kelvin(----)

```

```{r verifyfunction2-solution}

# Create vector `test` here.
test <- c(TRUE)

# celsius_to_kelvin() function: 

celsius_to_kelvin <- function(celsius) {
  stopifnot(is.numeric(celsius))
  kelvin <- celsius + 273.15
  return(kelvin)
}

# Input `test` into the celsius_to_kelvin() function. 

celsius_to_kelvin(test)

```

```{r verifyfunction2-check}
grade_code()
```

We see that R presents us with the error `Error in celsius_to_kelvin(n) : is.numeric(celsius) is not TRUE`. Note that had we not deployed the stopifnot() function, R would have tried to make sense of our input - coercing the logical vector `TRUE` into the numerical vector `1`. It would then proceed to calculate the temperature in Kelvins by using `1 + 273.15`, yielding the answer `274.15`. 


Yet, we see that the error message that we receive isn't the most effective. Let's now use the `if()` and `stop()` functions in conjunction with one another, to create successful, and effective error messages. 

### Error messages with `if()` and `stop()`. 

Let's say we only want to convert temperatures from Celsius to Kelvins for temperatures above 18 Degrees Celsius. How do we do this?

Let's first establish a check for the class of vectors. As in the previous exercise, we want to only use numeric vector inputs for our `celsius` arguments. 


```{r verifyfunction3, exercise = TRUE}


# celsius_to_kelvin() function: 

celsius_to_kelvin1 <- function(celsius) {
  
  # Put the validity check inside the `if()` statement. 
  
  if(-------) {
    
  # Add the error message "Sorry, this function only works for numeric inputs!"
  # in the body of the `stop()` function.  
    
    stop(--------)
  }
  
  kelvin <- celsius + 273.15
  return(kelvin)
}
```

```{r verifyfunction3-hint-1}
"Wrap `!is.numeric(celsius)` in the `if()` function"
```

```{r verifyfunction3-hint-2}
"Add 'Sorry, this function only works for numeric inputs!' in the body of the `stop()` function"
```

```{r verifyfunction3-solution}


# celsius_to_kelvin() function: 

celsius_to_kelvin1 <- function(celsius) {
  
  # Put the validity check inside the `if()` statement. 
  
  if(!is.numeric(celsius)) {
    
  # Add the error message "Sorry, this function only works for numeric inputs!"
  # in the body of the `stop()` function.  
    
    stop('Sorry, this function only works for numeric inputs!')
  }
  
  kelvin <- celsius + 273.15
  return(kelvin)
}

```


```{r verifyfunction3-check}
grade_code()
```

Check the function with the text input "Hi". Run code to see error message. 

```{r verifyfunction4, exercise = TRUE}

celsius_to_kelvin1("Hi")

```

```{r verifyfunction4-check}
grade_code()
```

Incredible! Let's now put an additional condition on our code to verify that no input below 18 degrees Celsius is converted by our function to Kelvins. 


```{r verifyfunction5, exercise = TRUE}


# celsius_to_kelvin() function: 

celsius_to_kelvin2 <- function(celsius) {
  if(!is.numeric(celsius)) {
    stop('Sorry, this function only works for numeric inputs!')
  }
  
  # Add a condition in the next `if()` function to make sure that the temperatures that can 
  # be input into the function are always 18 degrees or greater. 
  
  if(-----) {
    
    # Use the custom error message 'Sorry, this function only works 
    # for inputs greater than 18 Degrees Celsius' in the body of the `stop()` function.
    
    stop(-------------)
  }
  
  kelvin <- celsius + 273.15
  return(kelvin)
}

```

```{r verifyfunction3-hint}
"Use `celsius < 18` in the `if()` function."
```


```{r verifyfunction5-solution}


# celsius_to_kelvin() function: 

celsius_to_kelvin2 <- function(celsius) {
  if(!is.numeric(celsius)) {
    stop('Sorry, this function only works for numeric inputs!')
  }
  
  # Add a condition in the next `if()` function to make sure that the temperatures that can 
  # be input into the function are always 18 degrees or greater. 
  
  if(celsius < 18) {
    
    # Use the custom error message 'Sorry, this function only works 
    # for inputs greater than 18 Degrees Celsius' in the body of the `stop()` function.
    
    stop('Sorry, this function only works for inputs greater than 18 Degrees Celsius')
  }
  
  kelvin <- celsius + 273.15
  return(kelvin)
}

```

```{r verifyfunction5-check}
grade_code()
```

Great! Now verify the function for an argument input of less than 18. 

```{r verifyfunction6, exercise = TRUE}

celsius_to_kelvin2 <- function(celsius) {
  if(!is.numeric(celsius)) {
    stop('Sorry, this function only works for numeric inputs!')
  }
  if(celsius < 18) {
    stop('Sorry, this function only works for inputs greater than 18 Degrees Celsius')
  }
  kelvin <- celsius + 273.15
  return(kelvin)
}


celsius_to_kelvin2(--)

```

```{r verifyfunction6-check}
grade_code()
```

Great work! You're all set to verify the arguments of your functions. 

What we’ve accomplished:

- We’ve written several functions.
- We can check the validity of their input arguments.

## More on Arguments (R Primer Arguments 3)


### Default Values and `...` :

In this part, we generalize our functions, learn more technical details about R  default values for some arguments.

Let's start off with a short quiz: 

```{r, quiz-8}

quiz(
  question("True or False? A function can only have one argument.",
             answer("False", correct = TRUE),
             answer("True"),
             answer("False, but the second argument needs to be set to a default value."),
             answer("Depends on the function."),
             allow_retry = TRUE,
             message = "A function can have more than one argument. We have the freedom to assign reasonable default values to each of the arguments, depending on what the function is to be used for.")
  
)

```


###  

Assigning the right values to the right arguments can make or break your code. 

```{r, echo = TRUE}
log(1, 2)
log(2, 1)
## Uh oh.
```

But how much do you know about assigning arguments? Take this quiz to find out.

### methods

```{r methods, echo = FALSE}
question("Which method(s) can you use to match values to arguments? (Check all that apply.)",
         answer("Complete names (e.g. na.rm = TRUE)", correct = TRUE),
         answer("Partial names (e.g. na = TRUE)", correct = TRUE),
         answer("Numbers (e.g. 2 = TRUE)"),
         answer("Position (e.g. TRUE)", correct = TRUE),
         allow_retry = TRUE, type = "multiple", 
         incorrect = "Did you check _every_ method that works?",
         correct = "Good job! R can match values to arguments in three ways, but some of these methods are more foolproof then others.")
```

### Not matching


Let's now see where setting default arguments can become important. In the following questions, we use a simple function called `foo()`. 

```{r, echo = TRUE}
foo <- function(aaa = 1, abb = 2, abc = 3) {
  c(aaa, abb, abc)
}
```

```{r default-args, echo = FALSE}
question("Now what will `foo()` return?",
         answer('[1] "aaa" "abb" "abc"'),
         answer('[1] 1 2 3', correct = TRUE, message = "An argument with a default value is optional, R will use the default value if you do not provide one."),
         answer("An error"),
         allow_retry = TRUE)
```

```{r, echo = TRUE}
foo <- function(aaa = 1, abb = 2, abc = 3) {
  c(aaa, abb, abc)
}
```

```{r position-1, echo = FALSE}
question("What will `foo(0)` return?",
         answer("[1] 1 2 3"),
         answer("[1] 0 2 3", correct = TRUE, message = "If you do not provide an argument name in your call, R will match your value to the first unmatched argument name."),
         answer("[1] 1 0 3"),
         answer("[1] 1 2 0"),
         answer("[1] 0 0 0"),
         answer("An error"),
         allow_retry = TRUE)
```

###  

```{r, echo = TRUE}
foo <- function(aaa = 1, abb = 2, abc = 3) {
  c(aaa, abb, abc)
}
```


```{r position-2, echo = FALSE}
question("What will `foo(0, 5)` return?",
         answer("[1] 1 2 3"),
         answer("[1] 0 5 3", correct = TRUE, message = "I think you have the hang of position matching."),
         answer("[1] 0 5 0"),
         answer("[1] 0 5 5"),
         answer("An error"),
         allow_retry = TRUE)
```

### Names

```{r, echo = TRUE}
foo <- function(aaa = 1, abb = 2, abc = 3) {
  c(aaa, abb, abc)
}
```

```{r complete-names, echo = FALSE}
question("What will `foo(aaa = 0, abc = 5)` return?",
         answer("[1] 1 2 3"),
         answer("[1] 0 5 3"),
         answer("[1] 0 5 0"),
         answer("[1] 0 2 5", correct = TRUE, message = "Using complete argument names is the easiest way to avoid an error (and to write comprehensible code). But what if you use a partial argument name?"),
         answer("An error"),
         allow_retry = TRUE)
```

### Partial names

```{r, echo = TRUE}
foo <- function(aaa = 1, abb = 2, abc = 3) {
  c(aaa, abb, abc)
}
```

```{r partial-names, echo = FALSE}
question("What will `foo(aa = 0)` return?",
         answer("[1] 1 2 3"),
         answer("[1] 0 2 3", correct = TRUE, message = "You do not need to spell out the full argument name, so long as you provide enough of the name to uniquely identify the argument. Here `aa` could only be the beginning of the `aaa` argument."),
         answer("[1] 0 0 0"),
         answer("An error"),
         allow_retry = TRUE)
```

###  

```{r, echo = TRUE}
foo <- function(aaa = 1, abb = 2, abc = 3) {
  c(aaa, abb, abc)
}
```

```{r partial-names-2, echo = FALSE}
question("What will `foo(a = 0)` return?",
         answer("[1] 1 2 3"),
         answer("[1] 0 2 3"),
         answer("[1] 0 0 0"),
         answer("An error", correct = TRUE, message = "Here `a` does not uniquely match an argument name: it is the beginning of all three argument names. R isn't psychic, so it won't guess which argument you meant. Instead R returns an error. It's the right thing to do because it lets you fix your code."),
         allow_retry = TRUE)
```

### Dots

Now for some dark magic:

```{r, echo = TRUE}
foodoo <- function(...) {
  c(...)
}
```

```{r dots-1, echo = FALSE}
question("What will `foodoo(1, 2, 3)` return?",
         answer("[1] 1 2 3", correct = TRUE, message = "`...` is a special mechanism for passing arguments. `...` will match any argument not otherwise matched and pass those arguments on as a group. This is a convenient way to pass a set of arguments through to a function in the function body."),
         answer("An error."),
         answer("Oh my goodness! You've broken R!"),
         allow_retry = TRUE)
```

###  

```{r}
foodoo <- function(...) {
  c(...)
}
```

```{r dots-2, echo = FALSE}
question("What will `foodoo(a = 1, b = 2, c = 3)` return?",
         answer("[1] 1 2 3"),
         answer("[1] 1 2 3 _but with the names a b c above the values_", correct = TRUE, message = "If you provide names to your values, `...` will capture those as well and pass them on. So here, our code essentially runs `c(a = 1, b = 2, c = 3)`."),
         answer("An error."),
         allow_retry = TRUE)
```

###  

Here is one last wrinkle. Suppose we rewrite `foodoo` slightly:

```{r}
foodoo <- function(a, ...) {
  c(...)
}
```

```{r dots-3, echo = FALSE}
question("What will `foodoo(a = 1, b = 2, c = 3)` return?",
         answer("[1] 1 2 3 with the names a b c above the values"),
          answer("[1] 2 3 with the names b c above the values", correct = TRUE, message = "Remember that `...` only captures _unmatched_ values. Now that `a` is a formal argument, `a = 1` will be matched to it, leaving `b = 2` and `c = 3` to be captured and passed on."),
         answer("An error."),
         allow_retry = TRUE)
```

Congratulations! You've finished the quiz and know a lot about how R parses arguments.


Let's make a new function of our own: This one finds the first argument raised to the power of second argument and prints the result in appropriate format. We'll call it power. 


```{r default, exercise = TRUE}
power <- function(x, y) {
  
# function to print x raised to the power y
  
result <- --- 
  
print(paste(x,"raised to the power", y, "is", result))
}

```

```{r default-hint}
"Use `x^y to calculate x raised to the power of y`."
```


```{r default-solution}
power <- function(x, y) {
  
# function to print x raised to the power y
  
result <- x^y 
  
print(paste(x,"raised to the power", y, "is", result))
}

```

```{r default-check}
grade_code()
```

Great. Notice that the function has two arguments - x and y. Test the function to calculate 2 raised to the power 5. 

```{r default2, exercise = TRUE}
power <- function(x, y) {
  result <- x^y 
  print(paste(x,"raised to the power", y, "is", result))
}

power(--,--)
```

```{r default2-solution}
power <- function(x, y) {
  result <- x^y 
  print(paste(x,"raised to the power", y, "is", result))
}

power(2,5)
```


```{r default2-check}
grade_code()
```


Let's say we only want to use the function to calculate the squared value of the input argument x. How do we go about changing the function to do that?

```{r default3, exercise = TRUE}
power <- function(x, y = --- ) {
  result <- x^y 
  print(paste(x,"raised to the power", y, "is", result))
}

# Now, call the function without specifying the y argument. 
# Calculate the Squared value of 515. 

power()
```

```{r default3-solution}
power <- function(x, y = 2) {
  result <- x^y 
  print(paste(x,"raised to the power", y, "is", result))
}

# Now, call the function without specifying the y argument. 
# Calculate the Squared value of 515. 

power(515)
```

```{r default3-check}
grade_code()
```

Now that we have set the y argument to a default value, we do not need to specify the value of the y argument every time we run the function.

## Control Flow (From RPrimer 4)

### `if` and `else`

### Control Flow

_Control flow_ refers to the order in which a function executes the statements in its body of code.

By default, R functions will execute each line of code in the body in order, and then return the result of the last line of code. But it doesn't have to be this way. You can write functions that run some code in some situations and other code in other situations. 

Let's learn how!

### if

Take a look at the code below. What is happening? 

```{r, echo = TRUE}
x <- 1
```

```{r, echo = TRUE}
x == 1
x > 1
log(x)
if (x == 1) log(x)
```

```{r, echo = TRUE}
if (x > 1) log(x)
```

Try to describe what `if` is doing in your own words. Your best guess is okay!

###  

```{r if-quiz, echo = FALSE}
question("Which describes `if`'s behavior? (Check all that apply).",
         answer("`if` takes a logical test and a piece of code. It runs the code _if_ the test returns TRUE.", correct = TRUE, message = "`if` is a way to run code only in certain cases. When you use `if`, you first pass it a logical test surrounded by parentheses and then a piece of code."),
         answer("`if` takes a logical test and a piece of code. It does not run the code _if_ the logical test returns FALSE.", correct = TRUE, message = "`if` will run the piece of code if the logical test returns TRUE and not run the code if the logical test returns FALSE."),
         answer("`if` returns the results of the code that appears in _parentheses_ behind `if`."),
         answer("`if` always executes all of the code that follows it.", message = "`if` will always execute the logical test that appears in parentheses behind `if`. However, `if` will only execute the code that appears after the logical test if the logical test returns TRUE."),
         allow_retry = TRUE)
```

### Data cleaner

Think you have it? Let's check. We'll use `if` to write a useful function.

Many data sets use their own symbols to represent missing values. For example, NOAA will often use -99 to represent missing values in weather data sets. Let's write a function that checks whether a value is -99, and if so replaces the value with NA, like this:

```{r echo = TRUE}
clean <- function(x) {
  if (x == -99) x <- NA
  x
}
```

```{r, echo = TRUE}
clean(1)
clean(-99)
```

###  

Here is a start. `clean()` takes an object and returns the object, but `clean()` is missing an important piece of code.

* Add an `if` statement to the beginning of `clean()`. Your statement should assign NA to x if x equals -99. Then click Submit Answer.

```{r clean-setup}
x <- -99
```

```{r clean, exercise = TRUE}
clean <- function(x) {
  # add if statement here
  x
}
```

```{r clean-hint}
"Don't forget to use == to check for equality."
```


```{r clean-solution}
clean <- function(x) {
  if (x == -99) x <- NA
  x
}
```

```{r clean-check}
"clean() replaces x with NA if x is -99, otherwise clean() returns x as is."
```

### else

Here is a second version of `clean()` that uses on a new command. Can you tell what `else` does?

```{r}
clean <- function(x) {
  if (x == -99) NA else x
}
```

* Run `clean()` with several different values, `22`, `-99`, `3`. What does `clean()` return in each case? Why?

```{r else-setup}
clean <- function(x) {
  if (x == -99) NA else x
}
```

```{r else, exercise = TRUE}
```

###  

```{r else-quiz, echo = FALSE}
question("Which describes `else`? (Check all that apply).",
         answer("R interprets `else` as a continuation of the `if` statement that immediately precedes it.", correct = TRUE, message = "`else` always follows an `if` statement (if not, `else` returns an error), and R always sees the combined `if` and `else` statements as a single statement."),
         answer("Unlike `if`, `else` does not take a logical test.", correct = TRUE),
         answer("Like `if`, `else` takes a piece of code.", correct = TRUE),
         answer("`else` executes the piece of code if the `if` condition returns FALSE.", correct = TRUE),
         answer("`else` does not execute the piece of code if the `if` condition returns TRUE.", correct = TRUE, message = "Think of a combined `if` `else` statement as a complete description: if the logical test at the start of the statement returns TRUE, R will run the code that follows `if` (but not the code that follows `else`). If the condition returns FALSE, R will run the code that follows `else` (but not the code that follows `if`)."),
         allow_retry = TRUE)
```

### A word about syntax

Although you can put `if` and `else` on the same line, you shouldn't because it is easy for readers to miss the trailing `else` when they scan the code. Also, placing `if` and `else` on the same line can make very long lines.

It would be more common to write our function like this:

```{r, echo = TRUE}
clean <- function(x) {
  if (x == -99) NA 
  else x
}
```

R parses the `if` and `else` lines as a single statement as long as `else` is the first thing that follows the `if` statement. As a result, R will return the result of the combined if else statement if it appears at the end of a function.

### \{\}

You can also pass `if` and `else` chunks of code surrounded by braces, `{}`. Braces group multiple lines of code into a single "piece." When you use braces in an if else statement, R will run (or not run) everything between the braces.

In this example, R will run all three lines that follow `else` whenever x does not equal -99.

```{r echo = TRUE}
clean <- function(x) {
  if (x == -99) {
    NA
  } else {
    x <- x^2
    x <- sqrt(x)
    x
  }
}
```

When you use braces, indent everything between the braces by two spaces to make your code more readable. And of course, you can use braces to organize your code even if you have a single line of code between the braces.

### If else quiz

SAS often saves missing values as `"."`.

* Write a function named `clean2` that takes a value named `x` and returns an NA if the value is `"."` (and returns the value of `x` otherwise).

```{r clean2-setup}
x <- "."
```

```{r clean2, exercise = TRUE, eval = TRUE}

```

```{r clean2-hint}
"clean2() should closely resemble clean()."
```

```{r clean2-solution}
clean2 <- function(x) {
  if (x == ".") NA 
  else x
}
```

```{r clean2-check}
grade_code("What if you'd like to check for both -99 and . in the same function?")
```

### else if

In that case, you can use `else` to chain together multiple `if` statements.

```{r eval = FALSE}
clean <- function(x) {
  if (x == -99) NA 
  else if (x == ".") NA
  else x
}
```

###  

This does the same thing as 

```{r eval = FALSE}
clean <- function(x) {
  if (x == -99) NA 
  else {
      if (x == ".") NA
      else x
  }
}
```

`clean()` will:

1. Check whether `x == -99`. If so `clean()` will return NA and skip the rest of the code. If not, `clean()` will...
1. Check whether `x == "."`. If so, `clean()` will return NA and skip the rest of the code. If not, `clean()` will...
1. Evaluate `x` and return its value.

`else if` is more readable than nested if else statements, especially if you use many `else if`s.

###  

You can use `else` to string together as many if statements as you like. R will treat the result as a single multi-part if else statement. 
Be thoughtful with the order. R will always evaluate the clauses in order, executing the code in the _first clause whose condition is true_ and ignoring every clause after that. 

### Your turn

* Write a function named `clean()` that uses `if`, `else`, and `else if` statements to replace the following four values with NA before returning x, `-99`, `"."`, `""`, `"NaN"`. Then click Submit Answer.

```{r clean3-setup}
x <- ""
```


```{r clean3, exercise = TRUE, eval = TRUE}

```

```{r clean3-solution, echo = FALSE}
clean <- function(x) {
  if (x == -99) NA 
  else if (x == ".") NA
  else if (x == "") NA
  else if (x == "NaN") NA
  else x
}
```

```{r clean3-check}
grade_code("You can add as many else if's after if as you like. However, you cannot add an else if clause after an else because R will interpret the else as the end of a complete if else statement.")
```

### Quiz

```{r eval = FALSE, echo = TRUE}
foo <- function(x) {
  if (x > 2) "a"
  else if (x < 2) "b"
  else if (x == 1) "c"
  else "d"
}
foo(1)
```

```{r else-if-quiz}
question("What will this code return?",
         answer('"a"', message = "R will not retunr a because the condition 1 > 2 is false."),
         answer('"b"', correct = TRUE, message = 'The condition 1 < 2 is true, so R will evaluate the code that follows it (i.e. "b") and then skip the remainder of the multi-part if statement.'),
         answer('"a" "b"', message = "The conditions 1 < 2 and 1 == 1 are both TRUE, but R will stop after the _first_ true condition."),
         allow_retry = TRUE)
```

### Quiz

```{r eval = FALSE, echo = TRUE}
clean <- function(x) {
  if (x == -99) NA 
  if (x == ".") NA
  if (x == "NaN") NA
  x
}
clean(-99)
```

```{r clean3-quiz}
quiz(question("What will this code return?",
         answer("NA", message = "Did you notice that the `if` statements are not linked by `else`? What difference does this make?"),
         answer("-99", correct = TRUE, message = "Since the `if` clauses are not linked by `else`, R treats them as separate statements. R checks each if statement. After the first statement it runs NA, but does not return this as the result of the function (because this is not the final statement in the function). R does not run NA for the next two if statements because their conditions are false. Then R reaches `x`, which is the last line and last statement in the function. R evaluates `x`, which equals -99, and returns -99 as the result of the function."),
         allow_retry = TRUE))
```

### Congratulations!

You now know how to use `if` and `else` in your code. Let's look at another way to control flow in R.

### `return()` and `stop()`

###  

You can tell R to stop executing a function early with 

* `return()`
* `stop()`, and
* `stopifnot()`

Each will work only in the context of a function (because they stop the function). You wouldn't run these directly at the command line, but they provide a powerful way to control the flow of your functions: They can make if else statements unnecessary and they can even make your code less buggy.

### `return()`

When R encounters `return()` it will stop executing the function that called `return()`. If you pass a value to `return()`, R will return that value when it stops executing the function. Let's see how it works.

```{r}
impatient_square <- function(x) {
  return(x)
  x^2
}
```

```{r return-quiz, echo = FALSE}
question("Which value will `impatient_square(2)` return?",
         answer("2", correct = TRUE, message = "R stops executing `impatient_square()` and returns `x` when R encounters `return(x)`. It doesn't matter that `return(x)` is not the last line of the function."),
         answer("4", message = "R stops executing `impatient_square()` when it encounters `return(x)`. As a result, R does not evaluate `x^2`."),
         allow_retry = TRUE)
```

### R with a python accent

If you are a python user, you might already use `return()` ...unnecessarily. In python, you explicitly tell each function what to return, e.g. 

```python
def mysquare(x):
  y = x * x
  return y
```

Translated to R this becomes:

```{r}
my_square <- function(x) {
  y <- sum(x) / length(x)
  return(y)
}
```

But in R, this `return()` is not needed. R functions automatically return the result of their last line of code. In R, you can save `return()` for unusual control flow.

### Using `return()`

Remember this function? It didn't work as expected because we forgot to link our if statements with `else`.

* Fix the function not by adding `else`, but by adding `return()` in the right places. Then Click Submit Answer.

```{r return, exercise = TRUE}
clean <- function(x) {
  if (x == -99) NA 
  if (x == ".") NA
  if (x == "NaN") NA
  x
}
```

```{r return-solution}
clean <- function(x) {
  if (x == -99) return(NA)
  if (x == ".") return(NA)
  if (x == "NaN") return(NA)
  x
}
```

```{r return-check}
"This version is slightly easier to read than a linked if else tree because you can think about each if clause separately. You can often avoid long, nested if else trees by using return() thoughtfully."
```

### NULL

`clean()` is a fairly useful function, but it does have one flaw. 

* What happens when `x = NULL`? Run the code and find out.

```{r null, exercise = TRUE}
clean <- function(x) {
  if (x == -99) return(NA)
  if (x == ".") return(NA)
  if (x == "NaN") return(NA)
  x
}
clean(NULL)
```

```{r null-solution}
clean <- function(x) {
  if (x == -99) return(NA)
  if (x == ".") return(NA)
  if (x == "NaN") return(NA)
  x
}
clean(NULL)
```

###  

`clean()` cannot handle `NULL` because `if` returns an error when it evaluates `NULL == -99`. And, unfortunately, the error message isn't very clear. This is the perfect case for `stop()`.

```{r echo = FALSE}
clean <- function(x) {
  if (x == -99) return(NA)
  if (x == ".") return(NA)
  if (x == "NaN") return(NA)
  x
}
```

```{r nullo, exercise = TRUE, echo = TRUE}
clean(NULL)
```

### `stop()`

`stop()` behaves like `return()`, but instead of returning a value, `stop()` returns an error, complete with a custom error message. Can you tell how it works?

```{r eval = FALSE}
immovable_square <- function(x) {
  stop("I refuse to proceed.")
  x^2
}
```

```{r stop-quiz, echo = FALSE}
question("What does `immovable_square(2)` return?",
         answer("4", message = 'R will stop executing `immovable_square()` and throw an error when it comes to `stop("I refuse to proceed.")`.'),
         answer("Error in immovable_square(2) : I refuse to proceed.", correct = TRUE, message = 'R will stop executing `immovable_square()` and throw an error when it comes to `stop("I refuse to proceed")`. The error message will look just like this answer; R inserts the message that you pass as a character string to `stop()`.'),
         allow_retry = TRUE)
```

### Use `stop()`

* Use `if` and `is.null()` to add a `stop()` call at the beginning of `clean()`. The command should return the error message `"x is NULL"` whenever x is NULL. 
* Then click Submit Answer.

```{r stop, exercise = TRUE}
clean <- function(x) {
  if (x == -99) return(NA)
  if (x == ".") return(NA)
  if (x == "NaN") return(NA)
  x
}
```

```{r stop-solution}
clean <- function(x) {
  if (is.null(x)) stop("x is NULL")
  if (x == -99) return(NA)
  if (x == ".") return(NA)
  if (x == "NaN") return(NA)
  x
}
```

```{r stop-check}
"Now `clean()` can handle NULL values in an intelligent way. This pattern (if + stop) is so common that R provides a shortcut for it, `stopifnot()`."
```

### `stopifnot()`

`stopifnot()` is a more readable substitute for statements that combine `if` and `stop()`. Can you guess how it works?

```{r stopifnot-quiz, echo = FALSE}
question('Which of these does the equivalent of `if (x < 0) stop("x is less than zero")`?',
         answer('`stopifnot(x < 0)`', message = "Good guess! But the function isn't called `stopif()`; it's called `stopifnot()`."),
         answer('`stopifnot(x >= 0)`', correct = TRUE, message = "`stopifnot()` stops if a condition is _not_ met, which is slightly different than `if` + `stop()`."),
         allow_retry = TRUE)
```

### differences

`stopifnot()` is different from `if` + `stop()` in a few important ways: 

1. Instead of checking whether a condition is met, `stopifnot()` checks whether a condition is _not_ met.
1. `stopifnot()` does not pass along a custom error message. Instead, `stopifnot()` always explains that the condition was not true:

```{r morezero, exercise = TRUE, echo = TRUE}
    x <- -1
    stopifnot(x >= 0)
```

Notice that the first argument of `stopifnot()` should always be a logical condition, the inverse of the condition it replaces in an `if` + `stop()` statement. 

You can include additional logical conditions for `stopifnot()` to check after the first. Separate each with a comma.

###  

Think you have it? 

* Try replacing the `if` + `stop()` statement in `clean()` with `stopifnot()`. Then click Submit Answer.

```{r stopifnot, exercise = TRUE, eval = TRUE}
clean <- function(x) {
  if (is.null(x)) stop("x is NULL")
  if (x == -99) return(NA)
  if (x == ".") return(NA)
  if (x == "NaN") return(NA)
  x
}
```

```{r stopifnot-hint}
"You can reverse the result of `is.null()` by placing an `!` in front of it: `!is.null()`."
```

```{r stopifnot-solution}
clean <- function(x) {
  stopifnot(!is.null(x))
  if (x == -99) return(NA)
  if (x == ".") return(NA)
  if (x == "NaN") return(NA)
  x
}
```

```{r stopifnot-check}
grade_code("`stopifnot()` is both very readable and very useful. Click Continue to see why.")
```

### Defensive programming with `stopifnot()`

You can save yourself debugging time by writing your functions to _fail fast_ with clear error messages. To do this, think about situations that will lead to errors and then check for them with `stopifnot()` at the beginning of your code.

```{r, echo = TRUE}
clean <- function(x) {
  stopifnot(!is.null(x), is.numeric(x), length(x) == 1)
  
  if (x == -99) return(NA)
  x
}
```

If things go wrong, `stopifnot()` will help you see what you need to fix as soon as you run your function. Compare this to what will happen if you do not use `stopifnot()`:

1. Your code will run until it triggers a (perhaps unhelpful) error message
1. Your code may not trigger an error message, but return an incorrect result that you will think is true. This would very bad.

###  

Congratulations! You know two techniques of control flow, how to:

1. Run specific code in specific cases
1. Stop execution early

In the Advanced Control Flow tutorial, you'll learn how to combine logical tests in an if statement as well as how to write if statements that work with vectors, which is a prerequisite if you want to write vectorized functions.

## Advanced Control Flow (From RPrimer 5)

### && and ||

###   

This tutorial extends the Control Flow tutorial, where you learned how to use `if`, `else`, `return()`, and `stop()`.

Here you will learn how to 

1. combine logical tests in an if statement 
1. write if statements that work with vectors, which is a prerequisite if you want to write vectorized functions.

###  

Here's what `clean()` looked like at the end of the Control Flow tutorial. Do you notice that all of the if statements have the same outcome? 

```{r eval = FALSE, echo = TRUE}
clean <- function(x) {
  stopifnot(!is.null(x))
  if (x == -99) return(NA)
  if (x == ".") return(NA)
  if (x == "NaN") return(NA)
  x
}
```

Let's use your knowledge of logical tests to trim them down to a single if statement. 

###   

* Write a logical test that returns TRUE when `x` is -99 OR `x` is "." (Let's ignore the "NaN" case to keep things simple). Then click Submit Answer.

```{r logic, exercise = TRUE, eval = TRUE}

```

```{r logic-hint}
"You can combine two logical tests in R with `&` (and) and `|` (or), e.g. x < 0 & x > 1."
```

```{r logic-solution}
x == -99 | x == "."
```

```{r logic-check}
grade_code("This is the correct way to combine logical tests in R, but it has some downsides when you use it in an if statement.")
```

### & and |

`&` and `|` are R's boolean operators for combining logical tests.

* `&` stands for "and" will return `TRUE` if _**both**_ tests return `TRUE` and will return `FALSE` otherwise. 
* `|` stands for "or" will return `TRUE` if _**one or both**_ tests returns `TRUE` and will return `FALSE` otherwise. 

So,

```{r ninetynine, echo = TRUE, exercise = TRUE}
x <- -99
x == -99 | x == "."
```

###  

However, it is bad practice to use `&` and `|` to combine logical tests within an `if` condition. Why? Because:

1. there is something better (as you'll see in a minute)
1. `&` and `|` tend to generate warning messages when used with `if`

###   

As R operators, both `&` and `|` are vectorized which means that you can use them with vectors. This is very useful.

```{r, echo = TRUE}
x <- c(-99, 0 , 1)
x == -99
x == "."
x == -99 | x == "."
```

###  

However, `if` conditions are not vectorized. `if` expects the logical test contained within its parentheses to return a **single** `TRUE` or `FALSE`. If the condition returns a vector of `TRUE` or `FALSE`s, `if` will use the first value and show a warning message.

```{r,echo = TRUE}
x <- c(-99, 0 , 1)
if (x == -99 | x == ".") NA
```

### && and ||

You can avoid this by always using `&&` and `||` within your `if` conditions. `&&` and `||` are lazy substitutes for `&` and `|`. They are lazy in two ways.

First, `&&` and `||` always return a single `TRUE` or `FALSE`. If you give `&&` or `||` vectors, they will compare only the first elements of the vectors---and they will not return a warning message.

```{r, echo = TRUE}
x <- c(-99, 0 , 1)
x == -99 || x == "."
```
    
### Use ||

Let's use this to our immediate advantage.

* Replace the two `if` statements below with a single statement that tests whether x is `-99` or `"."` without throwing error messages.

```{r or, exercise = TRUE, eval = TRUE}
clean <- function(x) {
  stopifnot(!is.null(x))
  if (x == -99) return(NA)
  if (x == ".") return(NA)
  x
}
```

```{r or-hint}
"Like |, || expects a _complete_ logical test on each side of ||."
```

```{r or-solution}
clean <- function(x) {
  stopifnot(!is.null(x))
  if (x == -99 || x == ".") return(NA)
  x
}
```

```{r or-check}
grade_code("Now lets see what happens if you use clean() with a vector of values.")
```

###  

```{r eval = FALSE}
clean <- function(x) {
  stopifnot(!is.null(x))
  if (x == -99 || x == ".") return(NA)
  x
}
```

```{r vector-quiz}
question("What will `clean(c(-99, 0, 1))` return?",
         answer("NA 0 1", message = "The logical test will only consider the first value of the vector x, which here contains c(-99, 0, 1). Since that value is -99, the logical test will return TRUE and R will execute the code `return(NA)`, which returns a single NA."),
         answer("NA _with_ a warning message", message = "You certainly would get a warning message if your logical test used |, which will pass a vector of TRUEs and FALSEs to if, triggering the message. However, || will only pass on a single TRUE or FALSE."),
         answer("NA _without_ a warning message", correct = TRUE),
         answer("an error"),
         allow_retry = TRUE)
```

### Computation

The most important reason to use `||` instead of `|` is that `||` saves unnecessary computation when possible. This is the second way that `&&` and `||` are lazy.

When possible, `&&` and `||` jump to the correct conclusion after evaluating the first of the two logical tests (not so with `&` and `|`). 

* `&&` will return `FALSE` if the test on the left returns `FALSE` (because the combined test would return `FALSE`). 
* `||` will return `TRUE` if the test on the left returns `TRUE` (because the combined test would return `TRUE`)  
  
In either case, `&&` and `||` will not evaluate the test on the right. Run the code below to see check for yourselves. 
    
```{r stopping, exercise = TRUE, eval = TRUE}
x <- -99
if (x == -99 || stop("if you evaluate this.")) "I didn't evaluate stop()."
```
 
###  

How could you use this?

Remember how this code returns an error because `if` cannot handle the result of `NULL == -99`? Run the code to see. 

```{r testing, exercise = TRUE}
clean <- function(x) {
  if (x == -99) return(NA)
  x
}
clean(NULL)
```

### Quiz

Suppose we redefine `clean()` like this. Run the code now:

```{r, echo = TRUE}
clean <- function(x) {
  if (is.null(x) || x == -99) return(NA)
  x
}
```

```{r null-quiz}
question("Now what will `clean(NULL)` return?",
         answer("An error", message = "clean() certainly would return an error if it evaluated `x == -99` (and collapsed the results into a single result to pass to `if`), but _does_ `clean()` evaluate `x == -99`?"),
         answer("NA", correct = TRUE, message = "|| first evaluates `is.null(x)`. `is.null(NULL)` returns `TRUE`, so `||` stops and returns `TRUE` to `if`. As a result, R never evaluates the code that leads to an error."),
         allow_retry = TRUE)
```

### Vectorized if

###   

Burried in the last section is an interesting question: what if you _do_ want `clean()` to work with vectors? i.e.

```{r echo = TRUE}
clean <- function(x) ifelse(x == -99, NA, x)
```

```{r echo = TRUE}
clean(c(-99, 0, 1))
```

That would be a handy way to clean whole columns of data. How could you do it?

###   

Compare these two functions (one should seem familiar). What is different?

```{r echo = TRUE}
clean <- function(x) {
  if (x == -99) NA else x
}
clean2 <- function(x) {
  ifelse(x == -99, NA, x)
}
```

```{r ifelse-quiz, echo = FALSE}
question("Make a guess: do these two functions do the same thing?",
         answer("Yes!", correct = TRUE, message = "Well, you are sort of correct..."),
         answer("No!", correct = TRUE, message = "Well, you are sort of correct..."),
         answer("Sort of...", correct = TRUE, message = "You are absolutely right."), 
         allow_retry = TRUE,
         type = "single")
```

### `ifelse()`

`ifelse()` is a function that replicates an if else statement. It takes three arguments: a logical test followed by two pieces of code. If the test returns `TRUE`, `ifelse()` will return the results of the first piece of code. If the test returns `FALSE`, `ifelse()` will return the results of the second piece of code. 

So `clean(-99)` and `clean2(-99)` both return `NA`.

```{r, echo = TRUE}
clean(-99)
clean2(-99)
```

###  

However, unlike `if` and `else`, `ifelse` is vectorized. As a result, you can pass `ifelse()` a vector of values and it will apply the implied if else statement separately to each element of the vector.

```{r echo = TRUE}
x <- c(-99, 0, 1)
ifelse(x == -99, NA, x)
```

`clean2()` inherits this vectorized property from `ifelse()`. 

```{r echo = TRUE}
clean2(c(-99, 0, 1))
```

Compare that to `clean()` (which is non-vectorized because it relies on `if` and `else`, which are non-vectorized).

```{r echo = TRUE}
clean(c(-99, 0, 1))
```

### `if_else()`

The dplyr package offers a slight improvement on `ifelse()` named `if_else()`. `if_else()` is faster than `ifelse()`, but it requires you to make sure that each case in the if else statement returns the same type of object. For example, the statement needs to return a real number (or a string, or a logical, etc.) _whether or not_ the condition is `TRUE`.

No big deal, right? Well kind of.

```{r ifelsing, echo = TRUE, exercise = TRUE}
x <- c(-99, 0, 1)
if_else(x == -99, NA, x)
```

### NA

What happened? Recall that data in R comes in six atomic types. 

```{r type-quiz, echo = FALSE}
question("What type of data is NA?",
         answer("double (i.e. real numbers)"),
         answer("integer"),
         answer("character"),
         answer("logical", correct = TRUE, message = "It turns out that if you type NA at the command line, R saves the result as a logical."),
         answer("complex"),
         answer("raw"),
         allow_retry = TRUE)
```

###  

It is true:

```{r echo = TRUE}
typeof(NA)
```

So when you write `if_else(x == -99, NA, x)`, `if_else()` returns a logical in the first case and a double (real number) in the second (assuming `x` is a real number). 

You can get around this mishap in two ways:

1. Stick to `ifelse()`
1. Use a NA that comes with a type

### Types of NA

You may not realize it, but R comes with five types of NA. They all appear as `NA` when printed, but they are each saved with a separate data type. These are:

```{r echo = TRUE}
NA # logical
NA_integer_ # integer
NA_real_ # double
NA_complex_ # complex
NA_character_ # character
```

You can fix `if_else()` by being precise about which NA to use (most other R functions will convert the type of NA without bothering you).

```{r echo = TRUE}
x <- c(-99, 0, 1)
if_else(x == -99, NA_real_, x)
```

### Use `if_else`

* Fix the `if_else()` statement of `clean2()` to work with real numbers. Then click Submoit Answer.

```{r cleans, exercise = TRUE}
clean2 <- function(x) {
  ifelse(x == -99, NA, x)
}
```

```{r cleans-solution}
clean2 <- function(x) {
  ifelse(x == -99, NA_real_, x)
}
```

```{r cleans-check}
"Notice that this version of `clean2()` will work with real numbers, but not other types of data like characters. What if you want `clean2()` to work with all types of data? That's simple: stick with `ifelse()`."
```

### Vectorized else if

###  

What if you want to write a vectorized version of a multi-part if else tree? Like the tree in this function:

```{r echo=TRUE}

clean <- function(x) {
  if (x == -99) NA 
  else if (x == ".") NA
  else if (x == "") NA
  else if (x == "NaN") NA
  else x
}

```

In this case, neither `ifelse()` or `if_else()` will do. Why? Because each can only handle a single if condition, but our tree has four.

### `case_when()`

You can vectorize multi-part if else statements with dplyr's `case_when()` function. Here is how you would use `case_when()` to rewrite our `foo()` function from the Control Flow tutorial.

Here is the masterpiece in its original form

```{r echo = TRUE }
foo <- function(x) {
  if (x > 2) "a"
  else if (x < 2) "b"
  else if (x == 1) "c"
  else "d"
}
```

And here it is with `case_when()`.

```{r echo=TRUE}
foo2 <- function(x) {
  case_when(
    x > 2  ~ "a",
    x < 2  ~ "b",
    x == 1 ~ "c",
    TRUE   ~ "d"
  )
}
```

###  

And here are our foos in action to prove that `foo2()` is vectorized.

```{r echo=TRUE}
x <- c(3, 2, 1)
foo(x)
foo2(x)
```

Notice that 

1. `case_when()` returns a single case for each element, the first case whose left hand side evaluates to `TRUE`
1. The left hand side of the last case evaluates to `TRUE` no matter what the value of `x` is (In fact, the left hand side _is_ `TRUE`). This is an easy way to add an `else` clause to the end of `case_when()`.

Now let's look at the unusual syntax of `case_when()`.

### `case_when()` syntax

```{r eval = FALSE, echo = TRUE}
foo2 <- function(x) {
  case_when(
    x > 2  ~ "a",
    x < 2  ~ "b",
    x == 1 ~ "c",
    TRUE   ~ "d"
  )
}
```

Each argument of `case_when()` is a pair that consists of a logical test on the left hand side and a piece of code on the right hand side. The two are _always_ separated by a `~`.

Like `if_else()`, `case_when()` expects each case to return the same type of output. So keep those NA types handy: `NA`, `NA_integer_`, `NA_real_`, `NA_complex_`, `NA_character_`.

### Final Challenge

* Rewrite the multi-part version of `clean()` to use `case_when()`, which will allow `clean()` to handle vectors. Retain each case. Assume where necessary that `clean()` will only work with real numbers. Then click Submit Answer.

```{r when, exercise = TRUE}

clean <- function(x) {
  if (x == -99) NA 
  else if (x == ".") NA
  else if (x == "") NA
  else if (x == "NaN") NA
  else x
}
```

```{r when-hint}
"Use NA's that have the right type."
```

```{r when-solution}
clean <- function(x) {
  case_when(
    x == -99 ~ NA_real_, 
    x == "." ~ NA_real_,
    x == "" ~ NA_real_,
    x == "NaN" ~ NA_real_,
    TRUE ~ x
  )
}
```

```{r when-check}
grade_code('And if you noticed that a vector of real numbers would never contain ".", "", and "Nan" because they are strings, you are of course right. Thanks for playing along with the charade.')
```

### Congratulations!

You've learned how to alter the control flow of your functions with:

* `if`
* `else`
* `return()`
* `stop()`
* `stopifnot()`
* `ifelse()`

Not only that, you tackled two advanced methods: dplyr's `if_else()` and dplyr's `case_when()`.


## Being proactive about NAs

While most of the built in R datasets (`midwest`, `gapminder` etc.) do not contain missing values, on most occasions you will find that your datasets have missing data values. Let's explore this further. 

```{r, quiz-9}

quiz(
  question("What does the `na.rm` argument do, if set to FALSE?",
             answer("It removes all input NA values and allows the function to compute the input object." ),
             answer("It adds NA values to the object, to stop the function from running."),
             answer("It does not remove NA values, stopping the function from successfully running", correct = TRUE),
             answer("All of the above"),
             allow_retry = TRUE,
             message = "Many built-in R functions have an na.rm = argument through which you can specify how you want to handle NAs. Typically the default value is na.rm = FALSE and typical default behavior is to either let NAs propagate or to raise an error."
    )
)

```

Let's now look at a new function. Instead of subtracting the maximum and minimum values of a numeric variable, it adds them. We call this function `add_max_min()`. 

```{r}
add_max_min <- function(x, probs = c(0, 1), na.rm = FALSE) {
  stopifnot(is.numeric(x))
  the_quantiles <- quantile(x, probs, na.rm = na.rm)
  max(the_quantiles) + min(the_quantiles)
}

add_max_min(midwest$popadults)
```

```{r, quiz-10}

quiz(
  question("Assuming we convert the 10th value of the object `midwest$popadults` to NA, using the following code: `midwest$popadults[10] <- NA`. What stops this function from adding the value of the two quantiles?",
           
             answer("Nothing. It will work." ),
             answer("The probability arguments are incorrect."),
             answer("`na.rm` is set to `FALSE`.", correct = TRUE),
             answer("All of the above"),
             allow_retry = TRUE,
             message = "Correct! Our function, `add_max_min()` simply will not operate in the presence of NAs unless na.rm = TRUE"
    ),
  
  question("Where should we set `na.rm = TRUE`?",
           
             answer("Replace `x` in `function()` with `na.rm = TRUE`." ),
             answer("As an argument of the `quantile()` function."),
             answer("Add a new line after ` max(the_quantiles) + min(the_quantiles)`. Write `na.rm = TRUE` here."),
             answer("As an argument of `add_max_min`.", correct = TRUE),
             allow_retry = TRUE,
             message = "Correct! We could hardwire `na.rm = TRUE` within the `quantile()` function too. However, it is dangerous to invert the default behavior of a well-known built-in function and to provide the user with no way to override this."
    )
  
)

```

<!-- Not following the order of the book here - ... arguments were incorporated in the previous section of the tutorial. -->


## Anonymous Functions

Let's test our knowledge of anonymous functions. 

```{r, quiz-11}

quiz(
  question("What are anonymous functions?",
             answer("A built-in function in R."),
             answer("Functions without arguments."),
             answer("Functions with more than three arguments."),
             answer("Functions without names.", correct = TRUE),
             allow_retry = TRUE, 
             message = "Correct! An anonymous function in R is one without a name. We can create a function “on the fly”, especially for use within `map*()` functions, without bothering to give it a name."
    ),
  
  question("Which of the following map functions is not using an anonymous function? ",
             answer("`mutate(new = map_dbl(start, increment()))`",
                    correct = TRUE),
             answer("`mutate(new = map_dbl(start, function(x) x + 1))`"),
             answer("`mutate`(new = map_dbl(start, function(x) x + 1))`"),
             answer("`mutate(new = map_dbl(start, ~ . + 1))`"),
             answer("`mutate(new = map_dbl(start, ~ (. + 1)))`"),
             allow_retry = TRUE
    ),
  
   question("Which of the following can be classified as an anonymous function? ",
             answer("`(function(x) x * 10)(10)`",
                    correct = TRUE),
             answer("`fun <- function(x) x * 10`"),
             answer("`function(x * 10)`"),
             answer("`x * 10`"),
             allow_retry = TRUE
    )
  
  
)

```

Let's look at the following dataframe to see anonymous functions in action: 

###

```{r, echo = TRUE}

# Create a dataset
df <- data.frame(
  col1 = c("element1", "element2"),
  col2 = c("element1", "element2"), 
  stringsAsFactors = FALSE
)
```

Use `lapply()` to apply the function `function(x) paste(x, "doing stuff")` on the dataframe `df` we've created. Two arguments are relevant when using `lapply()`. The first is `x`, which is a vector (atomic or list) or an expression object. In our case, it will be our dataframe `df`. The second argument is `FUN`, the function to be applied to each element of `x`. 

```{r anon, exercise = TRUE}

# lapply an anonymous function to the columns of the dataset

lapply(--, -------------)

```

```{r anon-hint}

"The first argument of the `lapply()` function will be the dataframe, `df`."
```

```{r anon-solution}

# lapply an anonymous function to the columns of the dataset

lapply(df, function(x) paste(x, "doing stuff"))

```


```{r anon-check}

grade_code()

```

Excellent!

Note that instead of defining the function with a name, we resorted to using the complete code for the function to be applied to the data frame. Anonymous functions are commonly used with `lapply()` or `map*()` functions in this way. 

```{r, quiz-12}

quiz(
  question("Which of the following are using anonymous functions?",
             answer("`lapply(mtcars, function(x) length(unique(x)))`"),
             answer("`filter(function(x) !is.numeric(x), mtcars)`"),
             answer("`integrate(function(x) sin(x) ^ 2, 0, pi)`"),
             answer("All of the above", correct = TRUE),
             allow_retry = TRUE, 
             message = "Correct! All of the functions used above are anonymous functions!"
    )
  
)

```


##  List columns and `map_*()` functions


```{r, quiz-13}

quiz(
  question("Which of the following is true about list-columns?",
             answer("List columns, unlike lists, can contain vectors as elements in the list.",
                    correct = TRUE),
             answer("Lists and list columns are synonyms."),
             answer("List columns can only contain values as elements in the list."),
             answer("All of the above."),
             allow_retry = TRUE),
  
  
  question("Which of the following map functions is not using an anonymous function? ",
             answer("`mutate(new = map_dbl(start, increment()))`",
                    correct = TRUE),
             answer("`mutate(new = map_dbl(start, function(x) x + 1))`"),
             answer("`mutate(new = map_dbl(start, ~ . + 1))`"),
             answer("`mutate(new = map_dbl(start, ~ (. + 1)))`"),
             allow_retry = TRUE), 
  
   question("What is true about an atomic vector? Check all that apply.",
             answer("Each element of the atomic vector has more than one value."),
             answer("Each element of the atomic vector can take the form of a list."),
             answer("Each element of the atomic vector has one value.", correct = TRUE),
             answer("if an atomic vector is a column in your dataset, each observation gets a single value", correct = TRUE),
             allow_retry = TRUE, 
           type = "multiple"), 
  
    question("What will the following function do: `tibble(new_col = list(1:19, 23:25, 3:5))`. Check all that apply.",
             answer("Create a column variable with values starting from 1, and ending at 25."),
             answer("Create a tibble with one variable and two observations"),
             answer("Create a tibble with one variable and one observation.", correct = TRUE),
             answer("Produce a list with three elements", correct = TRUE),
             allow_retry = TRUE,
            type = "multiple"),

     question("Which of the following is true about list columns? Check all that apply.",
             answer("We use `str_split()` to remove list columns."),
             answer("We use `mutate()` to create list columns.", correct = TRUE),
             answer("Each element of a list can only have one kind of data type."),
             answer("Each element of a list can only have multiple data types.", correct = TRUE),
             allow_retry = TRUE,
             type = "multiple")

)

```

### Let's practice!

Let's create a simple tibble. At a race at the Olympics, a racer ran 4 laps. The time it took the racer to run each lap , in minutes, is to be fed into the tibble column, `mins`. 

The times are as follow: `c(10.3, 14.56, 17.9, 8.34)`. 

Create a tibble using the `tibble()` function, with the column variable `mins`.


```{r, map1, exercise = TRUE}

tibble(--------------)

```


```{r map1-solution, echo = TRUE, eval = TRUE}

tibble(mins = c(10.3, 14.56, 17.9, 8.34))

```

```{r map1-check}

grade_code(correct = "Great work!")

```

Let's now use this data to practice using the `map*()` function. Because we're dealing with numeric values here, we're going to use the `map_dbl()` function. Let's say we want to convert the time for each lap into seconds. How do we go about doing that? 


Let's name our new column `seconds`. We will use `mutate()` in conjunction with  `map_dbl()` to create a new column. 

Remember, the `map_dbl()` function applies the same function to every row in a tibble. Let's first use an anonymous function to do our job for us. 

```{r, map2, exercise = TRUE}

tibble(mins = c(10.3, 14.56, 17.9, 8.34)) %>% 
  mutate(seconds = map_dbl(----, ------------))

```

```{r map2-hint}
"Use `function(x) x * 60`."

```


```{r, map2-solution, echo = TRUE, eval = TRUE}

tibble(mins = c(10.3, 14.56, 17.9, 8.34)) %>% 
  mutate(seconds = map_dbl(mins, function(x) x * 60))

```

```{r map2-check}

grade_code(correct = "You have successfully created a new column for the same time, in seconds. Note that this could have also been achieved using a named function. Let's see how we can do that.")
```

Create a simple function called `convert` to change time in minutes to seconds. 

```{r, map3, exercise = TRUE}

convert <- -------------

```


```{r, map3-solution}

# Use code from the previous exercise question
convert <- function(x) x * 60

```

```{r map3-check}

grade_code(correct = "Amazing!")
```

Now use this function in tandem with the `map_dbl()` function to convert times from seconds, to minutes. 

```{r make-map4, echo = FALSE}
convert <- function(x) x * 60

```

```{r map4, exercise = TRUE}

# Use the `convert()` function

tibble(mins = c(10.3, 14.56, 17.9, 8.34)) %>% 
  mutate(seconds = map_dbl(mins, ------))

```

```{r map4-hint}
"Use `convert` without the parantheses that normally follow a function. This is a standard rule for all `map*()` functions."

```

```{r, map4-solution}
# Use the `convert()` function

tibble(mins = c(10.3, 14.56, 17.9, 8.34)) %>% 
  mutate(seconds = map_dbl(mins, convert))
```

```{r map4-check}

grade_code(correct = "Amazing!")
```

You have successfully used the `map_dbl()` function for converting minutes to seconds for the racer's four laps. An alternative way to write an anonymous function for the same purpose would be by using the `purrr` package (from where we get `map_dbl()`). This second way uses the ~ operator and then uses . to represent the current element.

For example, instead of using `map_dbl(mins, function(x) x * 60))`, we could have used `map_dbl(mins, ~ (. * 60))`. 

Try this in the section below. You should obtain the same results. 

```{r map5, eval=TRUE, echo=TRUE, exercise = TRUE}

tibble(mins = c(10.3, 14.56, 17.9, 8.34)) %>% 
  mutate(seconds = map_dbl(mins, ~ (. * 60)))


```

For our last exercise, let's revert back to our `midwest` data. For each of the counties in the data, we are aware of the percentage of people that are living below the poverty line (`midwest$percbelowpoverty`).

We will use this to calculate the percentage of people above poverty line, for conties Wexford and Waupaca. 

```{r make-map6}

midwest_data <- ggplot2::midwest


```

```{r map6,eval=TRUE, echo = TRUE}

midwest_data <- midwest_data %>% 
  
  # Selecting for relevant variables
  
  select(county, percbelowpoverty) %>% 
  
  # Filtering for the relevant counties.
  
  filter(county %in% c("WEXFORD", "WAUPACA")) %>% 
  
  # grouping by county.
  
  group_by(county)
 
midwest_data

```

Now, use summarize to create a list column, `pov` for the `percbelowpoverty` variable. 

```{r make-map7}

midwest_data <- ggplot2::midwest


midwest_data <- midwest_data %>% 
  
  # Selecting for relevant variables
  
  select(county, percbelowpoverty) %>% 
  
  # Filtering for the relevant counties.
  
  filter(county %in% c("WEXFORD", "WAUPACA")) %>% 
  
  # grouping by county.
  
  group_by(county)
 
```

```{r map7, exercise = TRUE, eval = TRUE, echo = TRUE}

midwest_data <- midwest_data %>% 
   
  # Use summarize to create a list column, `pov` for the `percbelowpoverty` variable.
  
  ---------------------

midwest_data
```

```{r map7-solution}

midwest_data <- midwest_data %>% 
   
  # Use summarize to create a list column, `pov` for the `percbelowpoverty` variable.
  
  summarize(pov = list(c(percbelowpoverty)))

midwest_data

```


```{r map7-check}

grade_code("Great, you have created a list column!")

```

Now use `map_dbl()` to calculate the percentage of people above poverty, using an anonymous function. Name this variable `abovepov`. Use `100 - x` to calculate the remaining percentage. 

```{r make-map8}
midwest_data <- ggplot2::midwest


midwest_data <- midwest_data %>% 
  
  # Selecting for relevant variables
  
  select(county, percbelowpoverty) %>% 
  
  # Filtering for the relevant counties.
  
  filter(county %in% c("WEXFORD", "WAUPACA")) %>% 
  
  # Grouping by county.
  
  group_by(county) %>% 
  
  
 summarize(pov = list(c(percbelowpoverty)))


```


```{r map8, exercise = TRUE, echo = TRUE, eval=TRUE}

midwest_data %>% 
  mutate(abovpov = ------------------------------)

```

```{r map8-hint-1}
"Use `100 - x` to calculate the remaining percentage"
```

```{r map8-hint-2}
"Use `map_dbl(pov, ~ 100 - .)` to calculate the remaining percentage"
```


```{r map8-solution}

midwest_data %>% 
  mutate(abovpov = map_dbl(pov, ~ 100 - .))

```

```{r map8-check}

grade_code("Great, you've calculated the percentage of population above the poverty line in Wexford and Waupaca!")

```

Congratulations! This brings the tutorial to an end. You have mastered functions. 

##  Using built in functions with list columbs and `map_*()` functions. 

1) Let's now introduce you to a series of questions that help you get the hang of using list columns and `map_*()` functions in conjunction with built-in R functions. 

Create a tibble called `firstib`. Within the tibble, create a list column called `new_col` with values `345:500` and `693:1200`. 

```{r, A1, exercise = TRUE, echo=TRUE, eval=TRUE}

firstib <- tibble(---- = -----------------) 

```

```{r A1-hint-1}
"Use `new_col = list(345:500, 693:1200)` to create the list column"
```

```{r, A1-solution}

firstib <- tibble(new_col = list(345:500, 693:1200)) 
```

```{r A1-check}

grade_code("Great!")

```

Now, create a `mean_col` that uses `map_dbl()` to calculate the mean of the two elements of the list column. Type `firstib` to view the contents of the tibble. 

```{r, A2, exercise = TRUE, echo=TRUE, eval=TRUE}

firstib <- tibble(new_col = list(345:500, 693:1200)) %>% 
  mutate(mean_col = --------------))
  
------

```

```{r A2-hint-1}
"Use `(new_col, mean)` within `map_dbl()`  to create the mean column"
```

```{r, A2-solution}

firstib <- tibble(new_col = list(345:500, 693:1200)) %>% 
   mutate(mean_col = map_dbl(new_col, mean))

firstib 

```

```{r A2-check}

grade_code("Great work!")

```

Let's now use `map_dbl()` to map the maximum values of these two vector elements as a new column, `max_col` within the tibble. We'll be using the `max()` function for this purpose. The tibble has been created for you, for the purposes of this exercise. 


```{r, echo=FALSE}

firstib <- tibble(new_col = list(345:500, 693:1200)) %>% 
   mutate(mean_col = map_dbl(new_col, mean))

```

```{r, A3, exercise=TRUE, eval = TRUE, echo = TRUE}

firstib %>% 
  mutate(---- = ------------------)

```

```{r A3-hint-1}
"Use `(new_col, max)` within `map_dbl()`  to create the `max_col` column"
```

```{r, A3-solution}

firstib %>% 
  mutate(max_col = map_dbl(new_col, max))


```

```{r A3-check}

grade_code("Great work! You're getting the hang of this!")

```

2) Let's now use another tibble that we've created for you, called `sectib`. View it in the space below by running the code `sectib`, followed by `str()` to view its contents.  

```{r}

sectib <- tibble(list_col = list(c(1, 5, 7), 
                       5, 
                       c(10, 10, 11)))

```

```{r, B1, echo = TRUE, exercise = TRUE}


# Use `str()` with `sectib` to view the tibble. 

sectib %>% 
  ---()

```

Now use the `sum()` function to create the column `list_sum`. This column uses the `map_dbl()` on the `list_col` column. 

Then use `pull()` on `list_sum` to obtain the values of the `list_sum` column. 


```{r, B2, exercise=TRUE, eval = TRUE, echo = TRUE}

sectib %>% 
  mutate(----------------------------) %>% 
  pull(-----)

```

```{r B2-hint-1}
"Use `mutate()` to create the `list_sum` column."
```

```{r B2-hint-2}
"Use `(list_col, sum)` within `map_dbl()`  to create the `list_sum` column"
```

```{r, B2-solution}

sectib %>% 
  mutate(list_sum = map(list_col, sum)) %>% 
  pull(list_sum)


```

```{r B2-check}

grade_code()

```

## Loops

### repeat

###  

Many Data Science tasks require you to do the same thing over and over again. This is boring work---but not for your computer!

But how do you ask your computer to repeat a task? 

* The purrr package provides one way. I recommend that you check out the Iteration primer to learn how to use purrr and its map functions.

* This tutorial will show you another way. You will learn how to repeat tasks with R's loop functions. You know know enough to understand loops, and loops will expand your ability to write useful functions.

### Did you know?

Did you know that some languages, like C, C++ and python, use loops as an _all purpose_ programming tool. This approach will not work well in R.

R's loops are the right tool for a very specific job, but they lose out to other methods for other jobs. This is due to R's user-oriented design: most R functions implement their own pre-optimized loops behind the scenes in a lower level language. These built-in loops are much faster than any loop you could write at the command line. 

But don't let this drive you loopy! You'll learn when and when not to use loops in R if you read through to the end of _When to use loops_. But first, we have a more pressing topic:

What is a loop anyways?

### loops

R contains three types of loops

1. `repeat`
1. `while` 
1. `for`

The simplest of these are `repeat` loops.

### repeat 

`repeat` repeats an expression over and over again. To use `repeat`, type `repeat` without trailing parentheses. Then use braces to enclose one or more lines of code, e.g.

```{r eval = FALSE, echo = TRUE}
repeat {
  print("Hello")
}
```

```{r echo = FALSE}
cat("Hello\nHello\nHello\nHello\nHello\n...")
```

`repeat` will execute all of the code between the braces. Then it will execute it again. And again. And again...

###  

```{r repeat-quiz, echo = FALSE}
question("When will `repeat` stop executing the code?",
         answer("After five times.", message = "Repeat will repeat your code until the end of time (I shortened things in the example above). This isn't particularly useful, so you will almost always use repeat with..."),
         answer("Never, unless you or your computer intervene.", correct = TRUE, 
         message = "Repeat will repeat your code until the end of time. This isn't particularly useful, so you will almost always use repeat with..."))
```

### break

When an R loop encounters the `break` command, it exits the loop. This lets you schedule the end of a `repeat` loop.

* How many times will R repeat the code in the loop below? Make a prediction then Click Submit Answer to see if you are right.

```{r repeat, exercise = TRUE}

n <- 1
repeat {
  print(n)
  if (n == 5) break
  n <- n + 1
}
```

```{r repeat-solution}
n <- 1
repeat {
  print(n)
  if (n == 5) break
  n <- n + 1
}
```

```{r repeat-check}
grade_code("R repeats the print statement five times and then exits the loop. Notice how we created an n variable to use with the loop. The loop reaches the end because we increment n by one each time we run the code. What would happen if you left out n <- n + 1? Don't try! The loop would print 1 for a very long time.")
```

###   

Here is a good question: 

What happens to the value of `n` when you run the loop? 

`n` is defined as `1` _outside_ of the loop. Is `n` still `1` when the loop is finished?

```{r results='hide'}
n <- 1
repeat {
  print(n)
  if (n == 5) break
  n <- n + 1
}
n
```

```{r n-setup}
n <- 1
repeat {
  if (n == 5) break
  n <- n + 1
}
```

* Click Submit Answer to find out.

```{r n, exercise = TRUE}
n
```

```{r n-solution}
n
```

```{r n-check}
grade_code("n = 5. Are you surprised? R's loops are not like functions, which safely execute code in a new environment. R's loops execute code in the environment that calls them, which means that loops will modify the objects in the environment. This is important to know. It can be very useful---but not if you forget that it happens.")
```

### count

One way to prevent loops from altering objects in your environment is to put the loop into a function. For example, here is a function that counts from one to the number `x`. Do you see how it works?

```{r echo = TRUE}
count <- function(x) {
  n <- 1
  repeat {
    print(n)
    if (n == x) break
    n <- n + 1
  }
}
```

* Can you write something similar? Write a function named `count_down` that counts from `x` to one. Be sure to arrange for the loop to end! Then click Submit Answer.

```{r countdown, exercise = TRUE}
count_down <- function(x) {
  
  
  
  
  
}
```

```{r countdown-hint-1}
"Begin by setting n = x. In which direction should you increment n to get to one?"
```

```{r countdown-hint-2}
"Increment n by subtracting one from it on each repetition. You will want to do this _after_ you print the value of n for that repetition."
```

```{r countdown-hint-3}
"Arrange to break the loop after you print n when n = 1."
```

```{r countdown-solution}
count_down <- function(x) {
  n <- x
  repeat {
    print(n)
    if (n == 1) break
    n <- n - 1
  }
}
```

```{r countdown-check}
grade_code("You know, this sounds a lot like a rocket launch. Let's try to make the loop return the words 'Blast off!' when it ends. If we can do that, we could make loops return _anything_ when they end.")
```

### return

`break` is not the only way to end a loop in R. 

If your loop is in a function, you can end the loop with a `return()` statement. R will stop executing the function (and therefore the loop) and return the value supplied by `return()`. 

* Alter your function below to end the loop with `return("Blast off!")`. You will no longer need the `break` command. Then click Submit Answer.

```{r blastoff, exercise = TRUE}
count_down <- function(x) {
  n <- x
  repeat {
    print(n)
    if (n == 1) break
    n <- n - 1
  }
}
```

```{r blastoff-hint}
"Replace `break` with something else."
```

```{r blastoff-solution}
count_down <- function(x) {
  n <- x
  repeat {
    print(n)
    if (n == 1) return("Blast off!")
    n <- n - 1
  }
}
```

```{r blastoff-check}
grade_code("Return is a useful way to return a message or value from a loop within a function. And for satisfaction's sake...")
```

###  

```{r echo = TRUE}
count_down <- function(x) {
  n <- x
  repeat {
    print(n)
    if (n == 1) return("Blast off!")
    n <- n - 1
  }
}
```

```{r echo=TRUE}
count_down(10)
```

Congratulations! 

A loop executes a piece of code multiple times, perhaps allowing changes to accumulate as it goes.

###  

You've learned the essence of loops, now it is time to learn the forms. In the next sections, you will learn how to refine `repeat` loops into two common types of loops: 
1. `while` loops, and
1. `for` loops

Each of these is a `repeat` loop adapted to a specific task.

###  

Along the way, you will work through a simple project: you will write and then refine a function that checks whether a number is prime. (This is a common computer science right of passage).

Let's begin the project now with a `repeat` loop. You'll need to know some things about prime numbers and `%%` to begin...

###   Prime numbers

In mathematics, a _prime_ number is a number that can be divided evenly only by itself and the number one. In other words, if you try to divide a prime number by any number less than itself, you get a remainder (_to keep things simple, let's not worry about zero or negative numbers_). 

For example, five is a prime number because you get a remainder when you divide it by two, three, and four:

```{r results ='hold', echo = TRUE}
5 / 5
5 / 4
5 / 3
5 / 2
5 / 1
```

Interestingly, it is hard to prove that a number is prime...

###  

...unless you are a computer. Then you can divide the number by every number less than itself and show that the result has a remainder.

###  %%

To do this, you'll need to use the modulo function, `a %% b`, which we met in the control flow tutorial. Modulo is an arithmetic operator that returns the remainder of dividing `a` by `b`. For example, when you divide five by five, nothing is left over. When you divide five by four, one is left over. When you divide five by three, two is left over. And so on.

```{r results='hold', echo = TRUE}
5 %% 5
5 %% 4
5 %% 3
5 %% 2
5 %% 1
```

Modulo doesn't return a decimal, it returns the number that remains once you subtract the largest multiple of `b` from `a` (i.e the number that you would get for the remainder if you did long division).

### `is_prime`

The code below uses a `repeat` loop to provide the beginnings of a function. Complete the loop:

1. Change `n` to begin at `2` instead of `1`
1. Add an `if` statement that ends the loop and returns `TRUE` if/when `n == x`.
1. Add a second `if` statement that that checks whether `x %% n == 0` and returns `FALSE` if so.
1. Then Click Submit Answer

```{r prime-repeat, exercise = TRUE}
is_prime <- function(x) {
  n <- 1
  repeat {
    print(n)
    
    
    n <- n + 1
  }
}
```

```{r prime-repeat-solution}
is_prime <- function(x) {
  n <- 2
  repeat {
    print(n)
    if (n == x) return(TRUE)
    if (x %% n == 0) return(FALSE)
    n <- n + 1
  }
}
```

```{r prime-repeat-check}
grade_code("The function works because a non-prime number will trigger the second if statement and return FALSE before the loop progresses to n = x. A prime number will reach n = x and return TRUE.")
```

###  

* Use `is_prime` to check whether or not 89 is a prime number. Click Submit Answer to run your code.

```{r make-prime-repeat}
is_prime <- function(x) {
  n <- 2
  repeat {
    print(n)
    if (n == x) return(TRUE)
    if (x %% n == 0) return(FALSE)
    n <- n + 1
  }
}
```

```{r eighty-nine, exercise = TRUE, exercise.setup = "make-prime-repeat", echo = TRUE}

```

```{r eighty-nine-solution}
is_prime(89)
```

```{r eighty-nine-check}
grade_code("89 *is* prime. Now that you have a working prime function, let's modify it to use a more streamlined loop.")
```

### `while`

###  

`while` loops run while a logical condition is true. 

And, hey! So does our `repeat` loop! It runs _while_ n is less than x. How do you know? Because it stops and returns `TRUE` when n equals x.

```{r eval = FALSE, echo = TRUE}
is_prime <- function(x) {
  n <- 2
  repeat {
    print(n)
    if (n == x) return(TRUE)
    if (x %% n == 0) return(FALSE)
    n <- n + 1
  }
}
```

Let's replace `repeat` with `while`.

###  

`while` takes a logical test, like `if`, and a chunk of code, like `repeat`. `while` will repeat the code until the logical test returns `FALSE` (unless something in the code ends the loop first). 

```{r echo = TRUE}
n <- 1
while (n <= 5) {
  print(n)
  n <- n + 1
}
```

You can read `while` as "while this condition is true, repeat that."

### Can you use while?

* Rewrite `is_prime` to use `while` instead of `repeat`. The new loop should run while n is less than x. 
* Have `is_prime` return `TRUE` in its last line, after the loop is run. 
    * Under which conditions, will the function return `TRUE`?
* Then click Submit Answer.

```{r prime-while, exercise = TRUE}
is_prime <- function(x) {
  n <- 2
  repeat {
    print(n)
    if (n == x) return(TRUE)
    if (x %% n == 0) return(FALSE)
    n <- n + 1
  }
}
```

```{r prime-while-hint}
"The new loop should run while n is less than x."
```

```{r prime-while-solution}
is_prime <- function(x) {
  n <- 2
  while (n < x) {
    print(n)
    if (x %% n == 0) return(FALSE)
    n <- n + 1
  }
  TRUE
}
```

```{r prime-while-check}
grade_code("`is_prime` will return TRUE if the loop comes to a natural end. That will only happen if `x` is a prime number.")
```

###  

Congratulations! Your `while` loop works.

```{r echo = FALSE}
is_prime <- function(x) {
  n <- 2
  while (n < x) {
    print(n)
    if (x %% n == 0) return(FALSE)
    n <- n + 1
  }
  TRUE
}
```

```{r echo = TRUE}
is_prime(7)
```

In fact, it does the same thing as your `repeat` loop. So why use `while`? 

###  

`while` loops can be easier to write than `repeat` loops; you do not need to insert your own `if` statement. 

`while` loops are also easier to read: `while` makes the code author's intentions obvious.

Now let's look at `for`. `for` loops provide the same advantages for a different situation. A for loop says "for each of these values do this." 

### `for`

###  

A for loop takes a defined set of values and repeats the loop once for each of the values. 

And, hey! So does our while loop! It runs once for each value in the set {2, 3, 4,  ...x}. How do you know? Because the loop begins with n = 2, increments n by one, and ends when n = x. 

```{r eval = FALSE, echo = TRUE}
is_prime <- function(x) {
  n <- 2
  while (n < x) {
    print(n)
    if (x %% n == 0) return(FALSE)
    n <- n + 1
  }
  TRUE
}
```

Let's replace `while` with `for`.


### `for` syntax

The syntax of `for` is similar to `while`, but instead of taking a logical test, `for` takes a three part statement:

1. An object that may, or may not, be used in the code chunk that follows `for`.
2. `in` (this never changes)
3. A vector of values to iterate over. On each run of the loop, `for` will assign a different value of the vector to the object named in step 1.


```{r echo = TRUE}
for (n in c(1, 2, 3, 4, 5)) {
  print(n)
}
```

`for` automatically ends the loop once it has iterated over each value in the vector.

### for example

`for` loops are very versatile. For example, `for` makes it easy to increment over strange sets of values:

```{r echo = TRUE}
for (i in c(1, 10, 200)) {
  print(i)
}
```

### for the win

You can also increment over "un-ordered" sets. You don't even need to stick to numbers:

```{r echo = TRUE}
for (person in c("Betty", "Veronica", "Archie")) {
  greeting <- paste("Hello", person)
  print(greeting)
}
```

### `seq`

That's great! But how can you pass `for` the set {2, 3, 4,  ...x}? You'll need to do something like that to retool `is_prime`.

One possibility is to use `seq(2, x)`, which creates a sequence of integers from 2 to x.

```{r echo = TRUE}
seq(1, 10)
```

`seq`, `seq_len` and `seq_along` form a complete family of helper functions for creating sequences in R.

### Can you use `for`?

* Convert `is_prime` to use a for loop that iterates over `n in seq(2, x - 1)`. 
    * Why wouldn't you want the loop to go all the way to n = x?
* Then click Submit Answer.

```{r for-1, exercise = TRUE}
is_prime <- function(x) {
  n <- 2
  while (n < x) {
    print(n)
    if (x %% n == 0) return(FALSE)
    n <- n + 1
  }
  TRUE
}
```

```{r for-1-hint}
"You no longer need to define n with an initial value, nor do you need to increment n."
```

```{r for-1-solution}
is_prime <- function(x) {
  for (n in seq(2, x - 1)) {
    print(n)
    if (x %% n == 0) return(FALSE)
  }
  TRUE
}
```

```{r for-1-check}
grade_code("You don't want the loop to continue to the case where n = x because x %% x always returns zero, even if x is prime, which means our function would always return FALSE. That wouldn't be good.")
```

###  

Here's an odd question: what if you wanted to skip some value in a loop?

For example, what if  we wanted to skip n = 5 when we run our loop?

You can do that with the last loop helper provided by R, `next`.

### `next`

When R encounters `next` in a loop, it will move on to the next iteration of the loop without executing the rest of the loop. Here, when n = 5, `next` causes R to move on to the next iteration of the loop (where n will equal 6).

```{r eval = FALSE}
is_prime <- function(x) {
  for (n in seq(2, x - 1)) {
    if (n == 5) next
    print(n)
    if (x %% n == 0) return(FALSE)
  }
  TRUE
}
```

If five is the last value in the loop, `next` will cause R to exit the loop.

### Quiz

```{r eval = FALSE, echo = TRUE}
n <- 1
while (n < 2) {
  print(n)
  next
  n <- n + 1
}
```

```{r next-quiz, echo = FALSE}
question("Will this loop run forever?",
         answer("No"),
         answer("Yes", correct = TRUE, message = "n will never increment to two  because R will skip over the end of the loop each time it encounters next. Be thoughtful when you write loops!"),
         allow_retry = TRUE)
```

###  

Congratulations! You've learned how to use all of R's loop functions! To be a loop master, click on the button to learn when to use loops in R and when not to.

### When to use loops

###  

I mentioned earlier that loops should not appear as frequently in your R code as they would in your C, C++, or python code. Why is this? Because R is an extremely vectorized language.

### Vectorization

By "vectorized" I mean that most R functions are designed to take vectors as input and to return vectors as output. So if, for example, you'd like to take the square root of every value in a vector, you do not need to loop over the values in the vector...

```{r echo = TRUE}
x <- c(1, 2, 3, 4, 5)
y <- double(length = 5)
for (i in seq_along(x)) {
  y[i] <- sqrt(i)
}
y
```

...you can just pass the vector to the square root function.

```{r echo = TRUE}
sqrt(x)
```

And the best part about R's vectorization is that it is very fast!

###  

R's arithmetic operators are also vectorized. If you add two vectors, R will add each pair of elements in the vectors and return the results as a new vector.

```{r echo = TRUE}
c(1, 2, 3, 4, 5) + c(1, 2, 3, 4, 5)
```

###  

And if you use R functions to build your functions, then your functions will inherit R's vectorization:

```{r echo = TRUE}
round_square <- function(vec) {
  sqs <- sqrt(vec)
  round(sqs, digits = 2)
}
round_square(x)
```

###   

Vectorization reduces the need to use loops. You can think of each vectorized function as implementing a loop for you.

### recursion

Some problems cannot be solved with vectorized functions, but even for these you do not necessarily need to use a loop. 

For example, we could solve the prime problem with recursion. This is the strategy of having a function call itself. Until one of the recursively recalled functions returns an answer that is then passed up the call stack.

```{r echo=TRUE}
is_prime <- function(x, n = 2) {
  if (n == x) return(TRUE)
  else if (x %% n == 0) return(FALSE)
  else is_prime(x, n = n + 1)
}
```

```{r echo = TRUE}
is_prime(89)
is_prime(88)
```

### `map`

Or you could use a map function from the purrr package.

It is all up to you. The moral is that loops are not an all purpose tool in R. They are best reserved for problems that are not easy to solve with a vectorized function.

To learn more about when you definitely _should_ use a loop, check out Hadley Wickham's three suggestions in [Advanced R](http://adv-r.had.co.nz/Functionals.html#functionals-not) (be aware, it is advanced).
